---
title: "1.CalcHR_GeneratePoints"
output: html_document
date: "2024-02-22"
---


# Objectives

- Visualize the home range for each bat from Kernel density plots
- Visualize the home range for each bat from MCPs 
- Assign random points in those home ranges (n x 10 for example)

- Evaluate which approach to defining home ranges gives a better picture of how bats may actually use the space. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
output: html_document
```


```{r}
### Prepare the work space
library(SpatialKDE)
library(tmap)
library(tmaptools)
library(tidyverse)
library(sp)
library(sf)
library(rgeos)
library(adehabitatMA)
library(digest)
library(mime)
library(jsonlite)
library(crs)
library(dplyr)
library(adehabitatHR)
library(Rcpp)
library(raster)
library(rgdal)
library(munsell)
library(tibble)
library(pillar)
library(ggplot2)
library(ResourceSelection)
library(mapview)
library(amt)
library(sjPlot)
library(Rgb)
library(nloptr)
library(colorspace)
library(rlang)
library(lattice)
library(MASS)
library(dismo)
library(scales)
library(rJava)
library(beepr)
library(OpenStreetMap)
library(lubridate)
library(rgdal)
library(munsell)
library(pillar)
library(ResourceSelection)
library(mapview)
library(amt)
library(sjPlot)
library(Rgb)
library(nloptr)
library(colorspace)
library(rlang)
library(broom) 
library(MASS)
library(ggmap)
library(gridExtra)
library(raster)
library(leaflet)
library(terra)
library(KernSmooth)
library(RColorBrewer)
library(cowplot)
library(colorBlindness)
library(kableExtra)

getwd()
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/Documents/1. PhD_Main/GitHub_link/Nittedal/SpatialAndMorphology_MBRA_MMYS"

output <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs"

file.name <- "HomeRangeExploration"

todays_date <- Sys.Date()
 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
dir.name
 
output_today <- dir.name
output_today

dir.create(output_today)
output_today
# MBRA <- dt1 %>% filter(Species == "M.brandtii") %>% droplevels() # 7634 obs of 49 vars

```



## Import and prepare dataset 

```{r}
# Import the "onsite" observations of bats 
Bats <- read_delim("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Spatial and Morphology/Inputs/45%.csv",     
                     delim = ";", escape_double = FALSE, trim_ws = TRUE)

Bats45 <- Bats

str(Bats45)
dim(Bats45)
#1502 obs of 8 vars 
names(Bats45)
#[1] "Date"     "Time"     "BatID"    "Easting"  "Northing" "BD"       "Dir.obs"  "Gain"     

## Add column for species ID ##
mystacinus<- Bats45$BatID %in% c("Amelia","Marie", "Turid", "Reeda", "Daisy", "Line", "Stine", "Ethel", "Ragnhild", "Ida", "Louise", "Aricia")

Bats45$Species<-ifelse(mystacinus == TRUE, "MYSTACINUS", "BRANDTII")
Bats45$Species<- as.factor(Bats45$Species)

table(Bats45$Species, Bats45$BatID)

## Rename the Dir Obs column to replace the space with a period 
Bats45 <- Bats45 %>% rename(Dir.obs = "Dir obs")

## Factorize bat ID column as well 
Bats45$BatID <- as.factor(Bats45$BatID)

summary(Bats45$BatID)
#Amelia   Aricia   Astrid    Dagny    Daisy    Ethel      Ida     Kaja     Line   Louise    Maren    Marie     Nora   Phoebe Ragnhild    Reeda 
#16       90       81       83      134      117      119       91       55       64      141       31       49       62       28       70 
#Sofia   Steffi    Stine     Thea    Turid 
#39       39       84      101        8 

##The number of onsite plots per individual bat 

table(Bats45$Dir.obs, Bats45$BatID)
#Yes = onsite plots


M.mystacinus <- subset(Bats45, Species == "MYSTACINUS") %>% droplevels()
M.brandtii <- subset(Bats45, Species == "BRANDTII") %>% droplevels()

levels(M.mystacinus$BatID)
levels(M.brandtii$BatID)

par(mfrow=c(2,1))
plot(table(M.mystacinus$BatID), main = "Number of points Mmys")
plot(table(M.brandtii$BatID), main = "Number of points Mbra")

#Fix date/time

names(Bats45)[1:2]<-c("date.cap","time.cap")

Bats45$date.cap<-as.character(Bats45$date.cap)

Bats45$time.cap<-as.character(Bats45$time.cap)

#---note: looks like many records are missing date and time data (this is rectified further down in the script)

Bats45$datetime<-paste(Bats45$date.cap,Bats45$time.cap)

head(Bats45$datetime)

# slightly different approach to dateime than in previous analyses
Bats45$datetime.posix<- as_datetime(Bats45$datetime,format = "%d.%m.%Y %H:%M:%S", tz="CET")
summary(Bats45)


summary(Bats45$Species)
  # BRANDTII MYSTACINUS 
  #      686        816 

summary(Bats45$Dir.obs=="Yes"|Bats45$BD=="Yes")
## Observations when the bats were detected by a handheld heterodyne while tracking 
#   Mode   FALSE    TRUE 
# logical     772     730 

summary(Bats45$Dir.obs=="Yes") # Observations where the bat was seen 
#   Mode   FALSE    TRUE 
# logical    1040     462

summary(Bats45$BD=="Yes") # Observations where the bat was picked up by a handheld heterodyne 
#    Mode   FALSE    TRUE 
# logical     820     682 
```


### Decide on a color palette
```{r}

### Conisdering using dark light blue and dark blue for brandts and mystacinus 
batcolors <- c("#32a5ff", "#011f4b")
#                MBRA      MMY

displayColors(batcolors)

cvdPlot(replacePlotColor(
  displayColors(batbugc))) 

```


## Create spatial data frames 
```{r}
coords <- coordinates(Bats45)<-Bats45[,c("Easting","Northing")]
#assigning coordinates 

proj4string(Bats45)<-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
#Define the projection 
crs(Bats45)

Bats45<-Bats45[!is.na(Bats45$Easting) | !is.na(Bats45$Northing),]
#Remove NA values  - There were none to remove 


## Visualize to check if this worked 
sBats45 <- st_as_sf(Bats45, crs = "+proj=utm + zone=32 + datum=WGS84 + units=m + no_defs")
plot(sBats45)
mapview(sBats45) # This works! 
 
###############################################################################

#Run again to get per species data sets without NA's 
Mystacinus <- subset(Bats45, Species == "MYSTACINUS")
Brandtii <- subset(Bats45, Species == "BRANDTII")

### Format so that it is only the Bat ID for each species for home range size
MMYS <- M.mystacinus %>% dplyr::select(c(BatID, Easting, Northing)) %>% droplevels()

MBRA <- M.brandtii %>% dplyr::select(c(BatID, Easting, Northing)) %>% droplevels()

###############################################################################
# MMYS
###############################################################################

Mcoords <- coordinates(MMYS) <- MMYS[,c("Easting","Northing")]
#assigning coordinates 

proj4string(MMYS)<-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
#Define the projection 
crs(MMYS)

sMMYS <- st_as_sf(MMYS, crs = "+proj=utm + zone=32 + datum=WGS84 + units=m + no_defs")
plot(sMMYS )
mapview(sMMYS)

###############################################################################
# MBRA
###############################################################################
Bcoords <- coordinates(MBRA)<-MBRA[,c("Easting","Northing")]
#assigning coordinates 

proj4string(MBRA)<-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
#Define the projection 
crs(MBRA)

sMBRA <- st_as_sf(MBRA, crs = "+proj=utm + zone=32 + datum=WGS84 + units=m + no_defs")
plot(sMBRA )
mapview(sMBRA)

#writeOGR(MBRA, dsn = "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs", layer = "MBRApts", driver = "ESRI Shapefile")

#MBRAshp <- st_read("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs/MBRApts.shp")


#####################################################
MBRAll <- spTransform(MBRA, CRS("+proj=longlat"))
MMYSll <- spTransform(MMYS, CRS("+proj=longlat"))

leaflet() %>% addProviderTiles("CartoDB.VoyagerLabelsUnder") %>% 
  addCircleMarkers(data = MBRAll, radius = 3, color = "#32a5ff", fillOpacity = 0.5) %>% 
  addCircleMarkers(data = MMYSll, radius = 3, color = "#011f4b", fillOpacity = 0.5) 


```



## Create a 95% kernel for each individual bat 

## MBRA bats 
levels(MBRA$BatID)
"Astrid" "Dagny"  "Kaja"   "Maren"  "Nora"   "Phoebe" "Sofia"  "Steffi" "Thea"  

```{r}

bound_box <- bb(Bats45, ext = 1.5)

##################################
## Astrid 
Astrid <- subset(MBRA, BatID == "Astrid") 
Astrid@data$BatID <- droplevels(Astrid@data$BatID) # Remove the empty levels of other MBRA bats

str(MBRA)
str(Astrid)

kde.outputAstrid <- kernelUD(Astrid, h="href", grid = 1000)

plot(kde.outputAstrid)

kdeAstrid <- raster(kde.outputAstrid)

# sets projection to British National Grid
projection(kdeAstrid) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeAstrid <- mask(kdeAstrid, Astrid)

range95Astrid <- getverticeshr(kde.outputAstrid, percent = 95)

Astrid_sf <- st_as_sf(Astrid)
range95Astrid_sf<- st_as_sf(range95Astrid)


  tm_shape(Astrid_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill = "#32a5ff") +
tm_shape(range95Astrid_sf) + 
  tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill = "#32a5ff")


#################################################

## Dagny 
Dagny <- subset(MBRA, BatID == "Dagny") 
Dagny@data$BatID <- droplevels(Dagny@data$BatID) # Remove the empty levels of other MBRA bats

str(MBRA)
str(Dagny)

kde.outputDagny <- kernelUD(Dagny, h="href", grid = 1000)

plot(kde.outputDagny)

kdeDagny <- raster(kde.outputDagny)

# sets projection to British National Grid
projection(kdeDagny) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeDagny <- mask(kdeDagny, Dagny)

range95Dagny <- getverticeshr(kde.outputDagny, percent = 95)

Dagny_sf <- st_as_sf(Dagny)
range95Dagny_sf <- st_as_sf(range95Dagny)


  tm_shape(Dagny_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill = "#32a5ff") +
tm_shape(range95Dagny_sf) + 
  tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill = "#32a5ff") 

#################################################

## Kaja 
Kaja <- subset(MBRA, BatID == "Kaja") 
Kaja@data$BatID <- droplevels(Kaja@data$BatID) # Remove the empty levels of other MBRA bats

str(MBRA)
str(Kaja)

kde.outputKaja <- kernelUD(Kaja, h="href", grid = 1000)

plot(kde.outputKaja)

kdeKaja <- raster(kde.outputKaja)

# sets projection to British National Grid
projection(kdeKaja) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeKaja <- mask(kdeKaja, Kaja)

range95Kaja <- getverticeshr(kde.outputKaja, percent = 95)

Kaja_sf <- st_as_sf(Kaja)
range95Kaja_sf <- st_as_sf(range95Kaja)

  tm_shape(Kaja_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#32a5ff") +
tm_shape(range95Kaja_sf) + 
  tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#32a5ff") 

#################################################

## Maren 
Maren <- subset(MBRA, BatID == "Maren") 
Maren@data$BatID <- droplevels(Maren@data$BatID) # Remove the empty levels of other MBRA bats

str(MBRA)
str(Maren)

kde.outputMaren <- kernelUD(Maren, h="href", grid = 1000)

plot(kde.outputMaren)

kdeMaren <- raster(kde.outputMaren)

# sets projection to British National Grid
projection(kdeMaren) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeMaren <- mask(kdeMaren, Maren)

range95Maren <- getverticeshr(kde.outputMaren, percent = 95)

Maren_sf <- st_as_sf(Maren)
range95Maren_sf <- st_as_sf(range95Maren)

  tm_shape(Maren_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#32a5ff") +
tm_shape(range95Maren_sf) + 
  tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#32a5ff")

#################################################

## Nora 
Nora <- subset(MBRA, BatID == "Nora") 
Nora@data$BatID <- droplevels(Nora@data$BatID) # Remove the empty levels of other MBRA bats

str(MBRA)
str(Nora)

kde.outputNora <- kernelUD(Nora, h="href", grid = 1000)

plot(kde.outputNora)

kdeNora <- raster(kde.outputNora)

# sets projection to British National Grid
projection(kdeNora) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeNora <- mask(kdeNora, Nora)

range95Nora <- getverticeshr(kde.outputNora, percent = 95)

Nora_sf <- st_as_sf(Nora)
range95Nora_sf <- st_as_sf(range95Nora)

  tm_shape(Nora_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#32a5ff") +
tm_shape(range95Nora_sf) + 
  tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#32a5ff")


#################################################

## Phoebe 
Phoebe <- subset(MBRA, BatID == "Phoebe") 
Phoebe@data$BatID <- droplevels(Phoebe@data$BatID) # Remove the empty levels of other MBRA bats

str(MBRA)
str(Phoebe)

kde.outputPhoebe <- kernelUD(Phoebe, h="href", grid = 1000)

plot(kde.outputPhoebe)

kdePhoebe <- raster(kde.outputPhoebe)

# sets projection to British National Grid
projection(kdePhoebe) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdePhoebe <- mask(kdePhoebe, Phoebe)

range95Phoebe <- getverticeshr(kde.outputPhoebe, percent = 95)

Phoebe_sf <- st_as_sf(Phoebe)
range95Phoebe_sf <- st_as_sf(range95Phoebe)

  tm_shape(Phoebe_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#32a5ff") +
tm_shape(range95Phoebe_sf) + 
  tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#32a5ff") 


#################################################

## Sofia 
Sofia <- subset(MBRA, BatID == "Sofia") 
Sofia@data$BatID <- droplevels(Sofia@data$BatID) # Remove the empty levels of other MBRA bats

str(MBRA)
str(Sofia)

kde.outputSofia <- kernelUD(Sofia, h="href", grid = 1000)

plot(kde.outputSofia)

kdeSofia <- raster(kde.outputSofia)

# sets projection to British National Grid
projection(kdeSofia) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeSofia <- mask(kdeSofia, Sofia)

range95Sofia <- getverticeshr(kde.outputSofia, percent = 95)

Sofia_sf <- st_as_sf(Sofia)
range95Sofia_sf <- st_as_sf(range95Sofia)

  tm_shape(Sofia_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#32a5ff") +
tm_shape(range95Sofia_sf) + 
  tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#32a5ff")


#################################################

## Steffi 
Steffi <- subset(MBRA, BatID == "Steffi") 
Steffi@data$BatID <- droplevels(Steffi@data$BatID) # Remove the empty levels of other MBRA bats

str(MBRA)
str(Steffi)

kde.outputSteffi <- kernelUD(Steffi, h="href", grid = 1000)

plot(kde.outputSteffi)

kdeSteffi <- raster(kde.outputSteffi)

# sets projection to British National Grid
projection(kdeSteffi) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeSteffi <- mask(kdeSteffi, Steffi)

range95Steffi <- getverticeshr(kde.outputSteffi, percent = 95)

Steffi_sf <- st_as_sf(Steffi)
range95Steffi_sf <- st_as_sf(range95Steffi)

  tm_shape(Steffi_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#32a5ff") +
tm_shape(range95Steffi_sf) + 
  tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#32a5ff")

#################################################

## Thea 
Thea <- subset(MBRA, BatID == "Thea") 
Thea@data$BatID <- droplevels(Thea@data$BatID) # Remove the empty levels of other MBRA bats

str(MBRA)
str(Thea)

kde.outputThea <- kernelUD(Thea, h="href", grid = 1000)

plot(kde.outputThea)

kdeThea <- raster(kde.outputThea)

# sets projection to British National Grid
projection(kdeThea) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeThea <- mask(kdeThea, Thea)

range95Thea <- getverticeshr(kde.outputThea, percent = 95)

Thea_sf <- st_as_sf(Thea)
range95Thea_sf <- st_as_sf(range95Thea)

  tm_shape(Thea_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#32a5ff") +
tm_shape(range95Thea_sf) + 
  tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#32a5ff")

#######################################################
  

```

## MMYS bats

levels(MMYS@data$BatID)
 # [1] "Amelia"   "Aricia"   "Daisy"    "Ethel"    "Ida"      "Line"     "Louise"   "Marie"   
 # [9] "Ragnhild" "Reeda"    "Stine"    "Turid" 

```{r}

## Amelia 
Amelia <- subset(MMYS, BatID == "Amelia") 
Amelia@data$BatID <- droplevels(Amelia@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Amelia)

kde.outputAmelia <- kernelUD(Amelia, h="href", grid = 1000)

plot(kde.outputAmelia)

kdeAmelia <- raster(kde.outputAmelia)

# sets projection to British National Grid
projection(kdeAmelia) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeAmelia <- mask(kdeAmelia, Amelia)

range95Amelia <- getverticeshr(kde.outputAmelia, percent = 95)

Amelia_sf <- st_as_sf(Amelia)
range95Amelia_sf<- st_as_sf(range95Amelia)

  tm_shape(Amelia_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Amelia_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b")
# 
#################################################

## Aricia 
Aricia <- subset(MMYS, BatID == "Aricia") 
Aricia@data$BatID <- droplevels(Aricia@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Aricia)

kde.outputAricia <- kernelUD(Aricia, h="href", grid = 1000)

plot(kde.outputAricia)

kdeAricia <- raster(kde.outputAricia)

# sets projection to British National Grid
projection(kdeAricia) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeAricia <- mask(kdeAricia, Aricia)

range95Aricia <- getverticeshr(kde.outputAricia, percent = 95)

Aricia_sf <- st_as_sf(Aricia)
range95Aricia_sf<- st_as_sf(range95Aricia)

  tm_shape(Aricia_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Aricia_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b") 

#########################################################

## Daisy 
Daisy <- subset(MMYS, BatID == "Daisy") 
Daisy@data$BatID <- droplevels(Daisy@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Daisy)

kde.outputDaisy <- kernelUD(Daisy, h="href", grid = 1000)

kdeDaisy <- raster(kde.outputDaisy)

# sets projection to British National Grid
projection(kdeDaisy) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeDaisy <- mask(kdeDaisy, Daisy)

range95Daisy <- getverticeshr(kde.outputDaisy, percent = 95)

Daisy_sf <- st_as_sf(Daisy)
range95Daisy_sf<- st_as_sf(range95Daisy)

  tm_shape(Daisy_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Daisy_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b")
#########################################################

## Ethel 
Ethel <- subset(MMYS, BatID == "Ethel") 
Ethel@data$BatID <- droplevels(Ethel@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Ethel)

kde.outputEthel <- kernelUD(Ethel, h="href", grid = 1000)

kdeEthel <- raster(kde.outputEthel)

# sets projection to British National Grid
projection(kdeEthel) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeEthel <- mask(kdeEthel, Ethel)

range95Ethel <- getverticeshr(kde.outputEthel, percent = 95)

Ethel_sf <- st_as_sf(Ethel)
range95Ethel_sf<- st_as_sf(range95Ethel)

  tm_shape(Ethel_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Ethel_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b")


#########################################################

## Ida 
Ida <- subset(MMYS, BatID == "Ida") 
Ida@data$BatID <- droplevels(Ida@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Ida)

kde.outputIda <- kernelUD(Ida, h="href", grid = 1000)

plot(kde.outputIda)

kdeIda <- raster(kde.outputIda)

# sets projection to British National Grid
projection(kdeIda) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeIda <- mask(kdeIda, Ida)

range95Ida <- getverticeshr(kde.outputIda, percent = 95)

Ida_sf <- st_as_sf(Ida)
range95Ida_sf<- st_as_sf(range95Ida)

  tm_shape(Ida_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Ida_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b")

#########################################################

## Line 
Line <- subset(MMYS, BatID == "Line") 
Line@data$BatID <- droplevels(Line@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Line)

kde.outputLine <- kernelUD(Line, h="href", grid = 1000)

kdeLine <- raster(kde.outputLine)

# sets projection to British National Grid
projection(kdeLine) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeLine <- mask(kdeLine, Line)

range95Line <- getverticeshr(kde.outputLine, percent = 95)

Line_sf <- st_as_sf(Line)
range95Line_sf<- st_as_sf(range95Line)

  tm_shape(Line_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Line_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b")

#########################################################

## Louise 
Louise <- subset(MMYS, BatID == "Louise") 
Louise@data$BatID <- droplevels(Louise@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Louise)

kde.outputLouise <- kernelUD(Louise, h="href", grid = 1000)

kdeLouise <- raster(kde.outputLouise)

# sets projection to British National Grid
projection(kdeLouise) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeLouise <- mask(kdeLouise, Louise)

range95Louise <- getverticeshr(kde.outputLouise, percent = 95)

Louise_sf <- st_as_sf(Louise)
range95Louise_sf<- st_as_sf(range95Louise)

  tm_shape(Louise_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Louise_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b") 

#########################################################

## Marie 
Marie <- subset(MMYS, BatID == "Marie") 
Marie@data$BatID <- droplevels(Marie@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Marie)

kde.outputMarie <- kernelUD(Marie, h="href", grid = 1000)

kdeMarie <- raster(kde.outputMarie)

# sets projection to British National Grid
projection(kdeMarie) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeMarie <- mask(kdeMarie, Marie)

range95Marie <- getverticeshr(kde.outputMarie, percent = 95)

Marie_sf <- st_as_sf(Marie)
range95Marie_sf<- st_as_sf(range95Marie)

  tm_shape(Marie_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Marie_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b")
#########################################################

## Ragnhild 
Ragnhild <- subset(MMYS, BatID == "Ragnhild") 
Ragnhild@data$BatID <- droplevels(Ragnhild@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Ragnhild)

kde.outputRagnhild <- kernelUD(Ragnhild, h="href", grid = 1000)

plot(kde.outputRagnhild)

kdeRagnhild <- raster(kde.outputRagnhild)

# sets projection to British National Grid
projection(kdeRagnhild) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeRagnhild <- mask(kdeRagnhild, Ragnhild)


range95Ragnhild <- getverticeshr(kde.outputRagnhild, percent = 95)

Ragnhild_sf <- st_as_sf(Ragnhild)
range95Ragnhild_sf<- st_as_sf(range95Ragnhild)

  tm_shape(Ragnhild_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Ragnhild_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b")

  #########################################################

## Reeda 
Reeda <- subset(MMYS, BatID == "Reeda") 
Reeda@data$BatID <- droplevels(Reeda@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Reeda)

kde.outputReeda <- kernelUD(Reeda, h="href", grid = 1000)

kdeReeda <- raster(kde.outputReeda)

# sets projection to British National Grid
projection(kdeReeda) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeReeda <- mask(kdeReeda, Reeda)

range95Reeda <- getverticeshr(kde.outputReeda, percent = 95)

Reeda_sf <- st_as_sf(Reeda)
range95Reeda_sf<- st_as_sf(range95Reeda)

  tm_shape(Reeda_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Reeda_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b")

#########################################################

## Stine 
Stine <- subset(MMYS, BatID == "Stine") 
Stine@data$BatID <- droplevels(Stine@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Stine)

kde.outputStine <- kernelUD(Stine, h="href", grid = 1000)

plot(kde.outputStine)

kdeStine <- raster(kde.outputStine)

# sets projection to British National Grid
projection(kdeStine) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeStine <- mask(kdeStine, Stine)

range95Stine <- getverticeshr(kde.outputStine, percent = 95)

Stine_sf <- st_as_sf(Stine)
range95Stine_sf<- st_as_sf(range95Stine)

  tm_shape(Stine_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Stine_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b")

#########################################################

## Turid 
Turid <- subset(MMYS, BatID == "Turid") 
Turid@data$BatID <- droplevels(Turid@data$BatID) # Remove the empty levels of other MMYS bats

str(MMYS)
str(Turid)

kde.outputTurid <- kernelUD(Turid, h="href", grid = 1000)

plot(kde.outputTurid)

kdeTurid <- raster(kde.outputTurid)

# sets projection to British National Grid
projection(kdeTurid) <-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

masked_kdeTurid <- mask(kdeTurid, Turid)

range95Turid <- getverticeshr(kde.outputTurid, percent = 95)

Turid_sf <- st_as_sf(Turid)
range95Turid_sf<- st_as_sf(range95Turid)

  tm_shape(Turid_sf, bbox = bound_box) + 
  tmap_mode('view') +
  tm_basemap("CartoDB.Voyager") +
  tm_dots(fill= "#011f4b", size = .3, fill_alpha = .8) +
tm_shape(range95Turid_sf) + 
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill= "#011f4b")

```



## Assign random poins based on the 95% kernel densities 

```{r}
# Bats$BatID <- as.factor(Bats$BatID)
# 
# checkpoitns <- as.data.frame(summary(Bats$BatID))
# # List of the number of points for each bat used to calculate the kernels 
# 
# ## Brandts
# random.points.Astrid <- spsample(range95Astrid, 81*10, "random")
# mapview(random.points.Astrid)
# 
# random.points.Dagny <- spsample(range95Dagny, 83*10, "random")
# mapview(random.points.Dagny)
# 
# random.points.Kaja <- spsample(range95Kaja, 91*10, "random")
# mapview(random.points.Kaja)
# 
# random.points.Maren <- spsample(range95Maren, 141*10, "random")
# mapview(random.points.Maren)
# 
# random.points.Nora <- spsample(range95Nora, 49*10, "random")
# mapview(random.points.Nora)
# 
# random.points.Phoebe <- spsample(range95Phoebe, 62*10, "random")
# mapview(random.points.Phoebe)
# 
# random.points.Sofia <- spsample(range95Sofia, 39*10, "random")
# mapview(random.points.Sofia)
# 
# random.points.Steffi <- spsample(range95Steffi, 39*10, "random")
# mapview(random.points.Steffi)
# 
# random.points.Thea <- spsample(range95Thea, 101*10, "random")
# mapview(random.points.Thea)
# 
# 
# ## Mystacinus 
# random.points.Amelia <- spsample(range95Amelia, 16*10, "random")
# mapview(random.points.Amelia)
# 
# random.points.Aricia <- spsample(range95Aricia, 90*10, "random")
# mapview(random.points.Aricia)
# 
# random.points.Daisy <- spsample(range95Daisy, 134*10, "random")
# mapview(random.points.Daisy)
# 
# random.points.Ethel <- spsample(range95Ethel, 117*10, "random")
# mapview(random.points.Ethel)
# 
# random.points.Ida <- spsample(range95Ida, 119*10, "random")
# mapview(random.points.Ida)
# 
# random.points.Line <- spsample(range95Line, 55*10, "random")
# mapview(random.points.Line)
# 
# random.points.Louise <- spsample(range95Louise, 64*10, "random")
# mapview(random.points.Louise)
# 
# random.points.Marie <- spsample(range95Marie, 31*10, "random")
# mapview(random.points.Marie)
# 
# random.points.Ragnhild <- spsample(range95Ragnhild, 28*10, "random")
# mapview(random.points.Ragnhild)
# 
# random.points.Reeda <- spsample(range95Reeda, 70*10, "random")
# mapview(random.points.Reeda)
# 
# random.points.Stine <- spsample(range95Stine, 84*10, "random")
# mapview(random.points.Stine)
# 
# random.points.Turid <- spsample(range95Turid, 8*10, "random")
# mapview(random.points.Turid)
# 
# ### Create available datasets
# 
# ### M.brandtii 
# avail.Astrid.df <- data.frame(
#   used = rep(0, length(random.points.Astrid)),
#   Species = rep("M.brandtii", length(Astrid)),
#   BatID = rep("Astrid", length(Astrid)), 
#   coords = random.points.Astrid@coords) 
# 
# avail.Dagny.df <- data.frame(
#   used = rep(0, length(random.points.Dagny)),
#   Species = rep("M.brandtii", length(Dagny)),
#   BatID = rep("Dagny", length(Dagny)), 
#   coords = random.points.Dagny@coords)
# 
# avail.Kaja.df <- data.frame(
#   used = rep(0, length(random.points.Kaja)),
#   Species = rep("M.brandtii", length(Kaja)),
#   BatID = rep("Kaja", length(Kaja)), 
#   coords = random.points.Kaja@coords)
# 
# avail.Maren.df <- data.frame(
#   used = rep(0, length(random.points.Maren)),
#   Species = rep("M.brandtii", length(Maren)),
#   BatID = rep("Maren", length(Maren)), 
#   coords = random.points.Maren@coords)
# 
# avail.Nora.df <- data.frame(
#   used = rep(0, length(random.points.Nora)),
#   Species = rep("M.brandtii", length(Nora)),
#   BatID = rep("Nora", length(Nora)), 
#   coords = random.points.Nora@coords)
# 
# avail.Phoebe.df <- data.frame(
#   used = rep(0, length(random.points.Phoebe)),
#   Species = rep("M.brandtii", length(Phoebe)),
#   BatID = rep("Phoebe", length(Phoebe)), 
#   coords = random.points.Phoebe@coords)
# 
# avail.Sofia.df <- data.frame(
#   used = rep(0, length(random.points.Sofia)),
#   Species = rep("M.brandtii", length(Sofia)),
#   BatID = rep("Sofia", length(Sofia)), 
#   coords = random.points.Sofia@coords)
# 
# avail.Steffi.df <- data.frame(
#   used = rep(0, length(random.points.Steffi)),
#   Species = rep("M.brandtii", length(Steffi)),
#   BatID = rep("Steffi", length(Steffi)), 
#   coords = random.points.Steffi@coords)
# 
# avail.Thea.df <- data.frame(
#   used = rep(0, length(random.points.Thea)),
#   Species = rep("M.brandtii", length(Thea)),
#   BatID = rep("Thea", length(Thea)), 
#   coords = random.points.Thea@coords)
# 
# ### MMYS
# 
# avail.Amelia.df <- data.frame(
#   used = rep(0, length(random.points.Amelia)),
#   Species = rep("M.mystacinus", length(Amelia)),
#   BatID = rep("Amelia", length(Amelia)), 
#   coords = random.points.Amelia@coords)
# 
# avail.Aricia.df <- data.frame(
#   used = rep(0, length(random.points.Aricia)),
#   Species = rep("M.mystacinus", length(Aricia)),
#   BatID = rep("Aricia", length(Aricia)), 
#   coords = random.points.Aricia@coords)
# 
# avail.Daisy.df <- data.frame(
#   used = rep(0, length(random.points.Daisy)),
#   Species = rep("M.mystacinus", length(Daisy)),
#   BatID = rep("Daisy", length(Daisy)), 
#   coords = random.points.Daisy@coords)
# 
# avail.Ethel.df <- data.frame(
#   used = rep(0, length(random.points.Ethel)),
#   Species = rep("M.mystacinus", length(Ethel)),
#   BatID = rep("Ethel", length(Ethel)), 
#   coords = random.points.Ethel@coords)
# 
# avail.Ida.df <- data.frame(
#   used = rep(0, length(random.points.Ida)),
#   Species = rep("M.mystacinus", length(Ida)),
#   BatID = rep("Ida", length(Ida)), 
#   coords = random.points.Ida@coords)
# 
# avail.Line.df <- data.frame(
#   used = rep(0, length(random.points.Line)),
#   Species = rep("M.mystacinus", length(Line)),
#   BatID = rep("Line", length(Line)), 
#   coords = random.points.Line@coords)
# 
# avail.Louise.df <- data.frame(
#   used = rep(0, length(random.points.Louise)),
#   Species = rep("M.mystacinus", length(Louise)),
#   BatID = rep("Louise", length(Louise)), 
#   coords = random.points.Louise@coords)
# 
# avail.Marie.df <- data.frame(
#   used = rep(0, length(random.points.Marie)),
#   Species = rep("M.mystacinus", length(Marie)),
#   BatID = rep("Marie", length(Marie)), 
#   coords = random.points.Marie@coords)
# 
# avail.Ragnhild.df <- data.frame(
#   used = rep(0, length(random.points.Ragnhild)),
#   Species = rep("M.mystacinus", length(Ragnhild)),
#   BatID = rep("Ragnhild", length(Ragnhild)), 
#   coords = random.points.Ragnhild@coords)
# 
# avail.Reeda.df <- data.frame(
#   used = rep(0, length(random.points.Reeda)),
#   Species = rep("M.mystacinus", length(Reeda)),
#   BatID = rep("Reeda", length(Reeda)), 
#   coords = random.points.Reeda@coords)
# 
# avail.Stine.df <- data.frame(
#   used = rep(0, length(random.points.Stine)),
#   Species = rep("M.mystacinus", length(Stine)),
#   BatID = rep("Stine", length(Stine)), 
#   coords = random.points.Stine@coords)
# 
# avail.Turid.df <- data.frame(
#   used = rep(0, length(random.points.Turid)),
#   Species = rep("M.mystacinus", length(Turid)),
#   BatID = rep("Turid", length(Turid)), 
#   coords = random.points.Turid@coords)
# 
# ###
# #Combine used data for all bats
# Bats.used <- rbind(
#   Astrid, # brandts 
#   Dagny,
#   Kaja,
#   Maren,
#   Nora,
#   Phoebe,
#   Sofia,
#   Steffi,
#   Thea,
#   Turid,
#   Amelia, 
#   Aricia,  
#   Daisy, 
#   Ethel, 
#   Ida, 
#   Line, 
#   Louise, 
#   Marie, 
#   Ragnhild, 
#   Reeda, 
#   Stine)
# head(Bats.used)
# summary(Bats.used)
# dim(Bats.used)
# table(Bats.used$BatID)
# 
#   # Astrid    Dagny     Kaja    Maren     Nora   Phoebe    Sofia   Steffi     Thea    Turid   Amelia 
#   #     81       83       91      141       49       62       39       39      101        8       16 
#   # Aricia    Daisy    Ethel      Ida     Line   Louise    Marie Ragnhild    Reeda    Stine 
#   #     90      134      117      119       55       64       31       28       70       84 
# 
# #Combine available data for all Mbra individuals
# Bats.avail <- rbind(
#   avail.Astrid.df, # brandts 
#   avail.Dagny.df,
#   avail.Kaja.df,
#   avail.Maren.df,
#   avail.Nora.df,
#   avail.Phoebe.df,
#   avail.Sofia.df,
#   avail.Steffi.df,
#   avail.Thea.df,
#   avail.Turid.df, 
#   avail.Amelia.df, # mystacinus 
#   avail.Aricia.df,  
#   avail.Daisy.df, 
#   avail.Ethel.df, 
#   avail.Ida.df, 
#   avail.Line.df, 
#   avail.Louise.df, 
#   avail.Marie.df, 
#   avail.Ragnhild.df, 
#   avail.Reeda.df, 
#   avail.Stine.df)
# 
# head(Bats.avail)
# #   used    Species BatID coords.x coords.y
# # 1    0 M.brandtii  Thea 604263.7  6658850
# # 2    0 M.brandtii  Thea 607650.4  6659768
# # 3    0 M.brandtii  Thea 604509.3  6658959
# # 4    0 M.brandtii  Thea 608160.9  6659012
# # 5    0 M.brandtii  Thea 604377.8  6660165
# # 6    0 M.brandtii  Thea 604798.2  6658901
# 
# test <- Bats.avail %>% mutate(BatID = factor(BatID), 
#                              Species = factor(Species)) 
# summary(Bats45$Species)
# #   M.brandtii M.mystacinus
# #         686           816 
# summary(test$Species)
#   # M.brandtii M.mystacinus 
#   #       6860         8160  ## Good! 
# 
# test <- test %>% select(BatID, Species) %>% distinct() %>% arrange(Species)
#  print(test)
# #  1 Astrid   M. brandtii  
# #  2 Dagny    M. brandtii  
# #  3 Kaja     M. brandtii  
# #  4 Maren    M. brandtii  
# #  5 Nora     M. brandtii  
# #  6 Phoebe   M. brandtii  
# #  7 Steffi   M. brandtii  
# #  8 Thea     M. brandtii  
# #  9 Sofia    M. brandtii  
# # 10 Amelia   M. mystacinus
# # 11 Aricia   M. mystacinus
# # 12 Daisy    M. mystacinus
# # 13 Ethel    M. mystacinus
# # 14 Ida      M. mystacinus
# # 15 Line     M. mystacinus
# # 16 Louise   M. mystacinus
# # 17 Marie    M. mystacinus
# # 18 Ragnhild M. mystacinus
# # 19 Reeda    M. mystacinus
# # 20 Stine    M. mystacinus
# # 21 Turid    M. mystacinus
# 
# # test <- Bats %>% select(BatID, Species) %>% distinct() %>% arrange(Species)
# # print(test, n = 21)
# # 1 Astrid   M. brandtii  
# #  2 Dagny    M. brandtii  
# #  3 Kaja     M. brandtii  
# #  4 Maren    M. brandtii  
# #  5 Nora     M. brandtii  
# #  6 Phoebe   M. brandtii  
# #  7 Steffi   M. brandtii  
# #  8 Thea     M. brandtii  
# #  9 Sofia    M. brandtii  
# # 10 Amelia   M. mystacinus
# # 11 Aricia   M. mystacinus
# # 12 Daisy    M. mystacinus
# # 13 Ethel    M. mystacinus
# # 14 Ida      M. mystacinus
# # 15 Line     M. mystacinus
# # 16 Louise   M. mystacinus
# # 17 Marie    M. mystacinus
# # 18 Ragnhild M. mystacinus
# # 19 Reeda    M. mystacinus
# # 20 Stine    M. mystacinus
# # 21 Turid    M. mystacinus
# 
# Bats.avail.df <- Bats.avail %>% 
#   rename(Easting = coords.x, Northing = coords.y) 
# head(Bats.avail.df)
# #   used    Species BatID  Easting Northing
# # 1    0 M.brandtii  Thea 604263.7  6658850
# # 2    0 M.brandtii  Thea 607650.4  6659768
# # 3    0 M.brandtii  Thea 604509.3  6658959
# # 4    0 M.brandtii  Thea 608160.9  6659012
# # 5    0 M.brandtii  Thea 604377.8  6660165
# # 6    0 M.brandtii  Thea 604798.2  6658901
# names(Bats.avail.df)
# #"used"     "Species"  "BatID"    "Easting"  "Northing"
# dim(Bats.avail.df)
# #15020 5
# 
# Bats.used.df <- Bats.used@data
# Bats.used.df$used <- 1
# Bats.used.df$BatID <- as.character(Bats.used.df$BatID)
# head(Bats.used.df)
# names(Bats.used.df)
# #"BatID"    "Easting"  "Northing" "used"  
# dim(Bats.used.df)
# # 1502        4
# 
# # 1502     + 15020 = 16522
# 
# #Combine Used and Avilable data for all Mbra individuals
# rsf.data <- full_join(Bats.used.df, Bats.avail.df)
# dim(rsf.data)
# #16522     5 # good! 
# 
# rsf.data$BatID <- as.factor(rsf.data$BatID)
# summary(rsf.data$BatID)
#   # Amelia   Aricia   Astrid    Dagny    Daisy    Ethel      Ida     Kaja     Line   Louise    Maren 
#   #    176      990      891      913     1474     1287     1309     1001      605      704     1551 
#   #  Marie     Nora   Phoebe Ragnhild    Reeda    Sofia   Steffi    Stine     Thea    Turid 
#   #    341      539      682      308      770      429      429      924     1111       88 
# 
# ## Some NAs introduced in the species column that should be fixed.
# batmap <- Bats.avail %>% select(BatID, Species) %>% distinct()
# rsf.data1 <- rsf.data %>% select(-Species)
# rsf.data2 <- left_join(rsf.data1, batmap)
# rsf.data2$UTM <- "UTM32"
# 
# ## export as a simple dataframe
# #write.csv(rsf.data2, file = file.path(output_today, "used_available_pts_combined95KE.csv")) 
# 
# 
# ## Make into a spatial object that can also be exported 
# coords <- coordinates(rsf.data2)<-rsf.data2[,c("Easting","Northing")]
# #assigning coordinates 
# 
# proj4string(rsf.data2)<-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
# #Define the projection 
# crs(rsf.data2)
# 
# rsf.data2<-rsf.data2[!is.na(rsf.data2$Easting) | !is.na(rsf.data2$Northing),]
# #Remove NA values  - There were none to remove 
# 
# ## Visualize to check if this worked 
# srsf.data2 <- st_as_sf(rsf.data2, crs = "+proj=utm + zone=32 + datum=WGS84 + units=m + no_defs")
# mapview(srsf.data2) # This works! 
# output_today
# 
# ## Exort as a geopackage
# #st_write(srsf.data2, "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs/HomeRangeExploration_2023-09-28/bats_rsf_pts.gpkg", driver = "GPKG")

```


## Summary results for the manuscript 
```{r}
## Summary of the "raw" data 
# mystacinus<- Bats$BatID %in% c("Amelia","Marie", "Turid", "Reeda", "Daisy", "Line", "Stine", "Ethel", "Ragnhild", "Ida", "Louise", "Aricia")
# 
# Bats$Species<-ifelse(mystacinus == TRUE, "M. mystacinus", "M. brandtii")
# Bats$Species<- as.factor(Bats$Species)
# 
# B <- Bats %>% filter(Species == "M. brandtii") %>% droplevels()
# M <- Bats %>% filter(Species == "M. mystacinus") %>% droplevels()
# 
# levels(B$BatID)
# levels(M$BatID)
# 
# knitr::kable(table(B$BatID)) %>% kable_minimal() # Number of real obs 
# knitr::kable(table(M$BatID)) %>% kable_minimal() # Number of real obs 
# 
# summary(rsf.data2)
# rsf.data2$Species <- factor(rsf.data2$Species)
# rsf.data2$BatID <- factor(rsf.data2$BatID)
# 
# # rsf.data2B <- rsf.data2 %>% dplyr::filter(Species == "M.brandtii") 
# # rsf.data2M <- rsf.data2 %>% dplyr::filter(Species == "M.mystacinus") %>% droplevels()
# # 
# # knitr::kable(table(rsf.data2B $BatID)) %>% kable_minimal() # Number of total obs 
# # knitr::kable(table(rsf.data2M$BatID)) %>% kable_minimal() # Number of total obs 
# 
# knitr::kable(table(rsf.data2$BatID)) %>% kable_minimal() 
# # Number of total obs 
# 
# rsf.avail <- rsf.data %>% filter(used == 0)
# rsf.availB <- rsf.data %>% filter(Species == "M.brandtii") %>% droplevels()
# rsf.availM <- rsf.data %>% filter(Species == "M.mystacinus") %>% droplevels()
# 
# knitr::kable(table(rsf.availB $BatID)) %>% kable_minimal() # Number of avail obs 
# knitr::kable(table(rsf.availM$BatID)) %>% kable_minimal() # Number of avail obs 
# 
# ## Calculate the area of the home range 
# ## MBRA
# area <- area(range95Astrid)
# BatID <- "Astrid"
# Astrid_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Dagny)
# BatID <- "Dagny"
# Dagny_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Kaja)
# BatID <- "Kaja"
# Kaja_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Maren)
# BatID <- "Maren"
# Maren_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Nora)
# BatID <- "Nora"
# Nora_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Phoebe)
# BatID <- "Phoebe"
# Phoebe_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Sofia)
# BatID <- "Sofia"
# Sofia_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Steffi)
# BatID <- "Steffi"
# Steffi_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Thea)
# BatID <- "Thea"
# Thea_area <- as.data.frame(cbind(area, BatID ))
# 
# # MMYS
# area <- area(range95Amelia)
# BatID <- "Amelia"
# Amelia_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Aricia)
# BatID <- "Aricia"
# Aricia_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Daisy)
# BatID <- "Daisy"
# Daisy_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Ethel)
# BatID <- "Ethel"
# Ethel_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Ida)
# BatID <- "Ida"
# Ida_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Line)
# BatID <- "Line"
# Line_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Louise)
# BatID <- "Louise"
# Louise_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Marie)
# BatID <- "Marie"
# Marie_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Ragnhild)
# BatID <- "Ragnhild"
# Ragnhild_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Reeda)
# BatID <- "Reeda"
# Reeda_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Stine)
# BatID <- "Stine"
# Stine_area <- as.data.frame(cbind(area, BatID ))
# 
# area <- area(range95Turid)
# BatID <- "Turid"
# Turid_area <- as.data.frame(cbind(area, BatID ))
# 
# listarea <- list(Astrid_area, Dagny_area, Kaja_area, # M. brandtii 
#                  Maren_area, Nora_area, Phoebe_area, 
#                  Sofia_area, Steffi_area, Thea_area, 
#                  Amelia_area, Aricia_area, Daisy_area, # M. mystacinus
#                  Ethel_area, Ida_area, Line_area, 
#                  Louise_area, Marie_area, Ragnhild_area, 
#                  Reeda_area, Stine_area, Turid_area) #list of 21, good
# 
# ## The area column is in m2 - create another that is in km2 
# allarea <- listarea %>% reduce(full_join, by=c('BatID', 'area')) 
# allarea$area <- as.numeric(allarea$area)
# allarea$area_km <- round(allarea$area /1000000 ,2)
# 
# allarea <- allarea %>% 
#   mutate(Species = case_when(
#     BatID %in% c("Amelia","Marie", "Turid", 
#                  "Reeda", "Daisy", "Line", 
#                  "Stine", "Ethel", "Ragnhild", 
#                  "Ida", "Louise", "Aricia") ~ "M. mystacinus", 
#     TRUE ~ "M. brandtii")) %>% 
#   mutate(Species = factor(Species)) %>% select(-area)
# 
# allareaB <- allarea %>% filter(Species == "M. brandtii") %>% select(-Species) %>% droplevels()
# allareaM <- allarea %>% filter(Species == "M. mystacinus") %>% select(-Species) %>% droplevels()
# 
# knitr::kable(allareaB) %>% kable_minimal() # Number of avail obs 
# knitr::kable(allareaM) %>% kable_minimal() # Number of avail obs 
# 
# ## MBRA
# # mean points for brandts bats
# mean(c(81,
# 83,
# 91,
# 141,
# 49,
# 62,
# 39,
# 39,
# 101))
# 
# # mean total points for brandts bats 
# mean(c(891,
# 913,
# 1001,
# 1551,
# 539,
# 682,
# 429,
# 429,
# 1111))
# 
# #mean home range in km2
# mean(c(7.59,
# 1.94,
# 0.11,
# 1.60,
# 5.23,
# 12.74,
# 4.05,
# 1.90,
# 0.83))
# 
# ## MMYS
# # mean points for mystacinus bats
# mean(c(16,
# 90,
# 134,
# 117,
# 119,
# 55,
# 64,
# 31,
# 28,
# 70,
# 84,
# 8))
# 
# # mean total points for mystacincus bats 
# mean(c(176,
# 990,
# 1474,
# 1287,
# 1309,
# 605,
# 704,
# 341,
# 308,
# 770,
# 924,
# 88))
# 
# #mean home range in km2
# mean(c(0.46,
# 1.05,
# 0.39,
# 1.52,
# 0.57,
# 0.30,
# 0.61,
# 1.24,
# 0.40,
# 0.08,
# 1.87,
# 0.12))


```


# Figure 2 
With assistance from Sikt KI GPT-5.4 
```{r}
Mystacinus_points <- rbind(
  Amelia_sf,
  Aricia_sf,
  Daisy_sf,
  Ethel_sf,
  Ida_sf,
  Line_sf,
  Louise_sf, 
  Marie_sf,
  Ragnhild_sf,
  Reeda_sf,
  Stine_sf,
  Turid_sf
)

Mystacinus_ranges <- rbind(
  range95Amelia_sf,
  range95Aricia_sf,
  range95Daisy_sf,
  range95Ethel_sf,
  range95Ida_sf,
  range95Line_sf,
  range95Louise_sf, 
  range95Marie_sf,
  range95Ragnhild_sf,
  range95Reeda_sf,
  range95Stine_sf,
  range95Turid_sf
)

MystacinusHR <- 
  tm_shape(Mystacinus_ranges, bbox = bound_box) +
  tm_borders(col = "#011f4b", lwd = 2, alpha = 0.7) +
  tm_fill(col = "#011f4b", alpha = 0.1) +
  tm_shape(Mystacinus_points) +
  tm_dots(col = "#011f4b", size = 0.3, alpha = 0.8) +
  tm_basemap("CartoDB.Voyager") +
  tm_scalebar(position = c("top", "right"), width = 200, text.size = 1.5)

tmap_mode("view")
MystacinusHR


tmap_mode("view")

Brandts_points <- rbind(
  Astrid_sf,
  Dagny_sf,
  Kaja_sf,
  Maren_sf,
  Nora_sf,
  Phoebe_sf,
  Sofia_sf,
  Steffi_sf,
  Thea_sf
)

Brandts_ranges <- rbind(
  range95Astrid_sf,
  range95Dagny_sf,
  range95Kaja_sf,
  range95Maren_sf,
  range95Nora_sf,
  range95Phoebe_sf,
  range95Sofia_sf,
  range95Steffi_sf,
  range95Thea_sf
)



# transform all to same CRS
Brandts_points <- st_transform(Brandts_points, st_crs(Brandts_ranges))
Mystacinus_ranges <- st_transform(Mystacinus_ranges, st_crs(Brandts_ranges))
Mystacinus_points <- st_transform(Mystacinus_points, st_crs(Brandts_ranges))

# clean geometries
Brandts_ranges <- st_make_valid(Brandts_ranges[!st_is_empty(Brandts_ranges), ])
Brandts_points <- st_make_valid(Brandts_points[!st_is_empty(Brandts_points), ])
Mystacinus_ranges <- st_make_valid(Mystacinus_ranges[!st_is_empty(Mystacinus_ranges), ])
Mystacinus_points <- st_make_valid(Mystacinus_points[!st_is_empty(Mystacinus_points), ])

# shared extent across all layers
bb1 <- st_bbox(Brandts_ranges)
bb2 <- st_bbox(Brandts_points)
bb3 <- st_bbox(Mystacinus_ranges)
bb4 <- st_bbox(Mystacinus_points)

xmin <- min(bb1["xmin"], bb2["xmin"], bb3["xmin"], bb4["xmin"], na.rm = TRUE)
xmax <- max(bb1["xmax"], bb2["xmax"], bb3["xmax"], bb4["xmax"], na.rm = TRUE)
ymin <- min(bb1["ymin"], bb2["ymin"], bb3["ymin"], bb4["ymin"], na.rm = TRUE)
ymax <- max(bb1["ymax"], bb2["ymax"], bb3["ymax"], bb4["ymax"], na.rm = TRUE)

xpad <- (xmax - xmin) * 0.08
ypad <- (ymax - ymin) * 0.08

common_bbox <- st_bbox(c(
  xmin = xmin - xpad,
  xmax = xmax + xpad,
  ymin = ymin - ypad,
  ymax = ymax + ypad
), crs = st_crs(Brandts_ranges))

# Brandts
BrandtsHR <-
  tm_shape(Brandts_ranges, bbox = common_bbox) +
  tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) +
  tm_fill(fill_alpha=.1, fill = "#32a5ff") +
  tm_shape(Brandts_points) +
  tm_dots(fill = "#32a5ff", size = 0.3, fill_alpha = 0.8) +
  tm_basemap("CartoDB.Voyager") +
  tm_scalebar(position = c("top", "right"), width = 100, text.size = 0.6)

BrandtsHR

# Mystacinus 011f4b
MystacinusHR <-
  tm_shape(Mystacinus_ranges, bbox = common_bbox) +
  tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) +
  tm_fill(fill_alpha=.1, fill = "#011f4b") +
  tm_shape(Mystacinus_points) +
  tm_dots(fill = "#011f4b", size = 0.3, fill_alpha = 0.8) +
  tm_basemap("CartoDB.Voyager") +
  tm_scalebar(position = c("top", "right"), width = 100, text.size = 0.6)

MystacinusHR

# inset map of Norway
tmap_mode("plot")


## Study area inset 
study_center <- st_as_sf(
  data.frame(lon = 10.8985915, lat = 60.0587263),
  coords = c("lon", "lat"),
  crs = 4326
)

norway_bbox <- st_bbox(c(
  xmin = 4.0,
  xmax = 32.0,
  ymin = 57.5,
  ymax = 71.5
), crs = st_crs(4326))

norway_box_sf <- st_as_sfc(norway_bbox)

norway_inset <-
  tm_shape(norway_box_sf) +
  tm_basemap("CartoDB.Voyager") +
  tm_shape(study_center) +
  tm_symbols(
    col = "black",
    shape = 21,
    border.col = "black",
    fill = "white",
    size = 1.2
  ) +
  tm_scalebar(position = c("bottom", "right"), width = 10, text.size = 0.6) 

norway_inset


#

# save both at same size
tmap_save(
  BrandtsHR,
  filename = file.path(output_today, "BrandtsHR_map.tiff"),
  width = 8,
  height = 12,
  units = "in",
  dpi = 600
)

tmap_save(
  MystacinusHR,
  filename = file.path(output_today, "MystacinusHR_map.tiff"),
  width = 8,
  height = 12,
  units = "in",
  dpi = 600
)


tmap_save(
  norway_inset,
  filename = file.path(output_today, "NorwayInset.tiff"),
  width = 8,
  height = 12,
  units = "in",
  dpi = 600
)


# Map saved to C:\Users\apmc\OneDrive - Norwegian University of Life Sciences\1. Nittedal 2018-2020\Nittedal_Main\Analyses\Outputs\HomeRangeExploration_2026-06-08\MystacinusHR_map.tiff
# Resolution: 4800 by 7200 pixels
# Size: 8 by 12 inches (600 dpi)


## for supplementary 

tmap_mode("view")

CombinedHR <-
  tm_shape(Mystacinus_ranges, bbox = bound_box) +
  tm_borders(col = "#011f4b", lwd = 2, alpha = 0.7) +
  tm_fill(col = "#011f4b", alpha = 0.1) +
  
  tm_shape(Brandts_ranges) +
  tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
  tm_fill(fill_alpha=.1, fill = "#32a5ff") +
  
  tm_shape(Mystacinus_points) +
  tm_dots(col = "#011f4b", size = 0.3, alpha = 0.8) +
  
  tm_shape(Brandts_points) +
  tm_dots(col = "#32a5ff", size = 0.3, alpha = 0.8) +
  
  tm_basemap("CartoDB.Voyager") +
  tm_scalebar(position = c("top", "right"), width = 200, text.size = 1.5)

CombinedHR

tmap_save(
  CombinedHR,
  filename = file.path(output_today, "CombinedHR.tiff"),
  width = 8,
  height = 12,
  units = "in",
  dpi = 600
)

#### Further exploration 
#### Sattelite images and the three Brandtii that have the most influence on the RSF results 

# ## Now sattelite
# 
# CombinedHR <-
#   tm_shape(Mystacinus_ranges, bbox = common_bbox) +
#   tm_borders(col = "#011f4b", lwd = 2, alpha = 0.7) +
#   tm_fill(col = "#011f4b", alpha = 0.1) +
#   
#   tm_shape(Brandts_ranges) +
#   tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
#   tm_fill(fill_alpha=.1, fill = "#32a5ff") +
#   
#   tm_shape(Mystacinus_points) +
#   tm_dots(col = "#011f4b", size = 0.3, alpha = 0.8) +
#   
#   tm_shape(Brandts_points) +
#   tm_dots(col = "#32a5ff", size = 0.3, alpha = 0.8) +
#   
#   tm_basemap("Esri.WorldImagery") +
#   tm_scalebar(position = c("top", "right"), width = 200, text.size = 1.5)
# 
# CombinedHR
# 
# 
# 
# ### Now look at the most influential bats 
# 
# bigbats_points <- rbind(
#   Dagny_sf,
#   Maren_sf,
#   Thea_sf
# )
# 
# 
# bigbats_ranges <- rbind(
#   range95Dagny_sf,
#   range95Maren_sf,
#   range95Thea_sf
# 
# )
# 
# 
# # transform all to same CRS
# bigbats_points <- st_transform(bigbats_points, st_crs(bigbats_ranges))
# bigbats_ranges <- st_transform(bigbats_ranges, st_crs(bigbats_ranges))
# 
# 
# # clean geometries
# bigbats_ranges <- st_make_valid(bigbats_ranges[!st_is_empty(bigbats_ranges), ])
# bigbats_points <- st_make_valid(bigbats_points[!st_is_empty(bigbats_points), ])
# 
#   tm_shape(Mystacinus_ranges, bbox = common_bbox) +
#   tm_borders(fill_alpha=.7, col = "#011f4b", lwd = 2) +
#   tm_fill(fill_alpha=.1, fill = "#011f4b") +
#   tm_shape(Mystacinus_points) +
#   tm_dots(fill = "#011f4b", size = 0.3, fill_alpha = 0.8) +
#   tm_basemap("CartoDB.Voyager") +
#   tm_scalebar(position = c("top", "right"), width = 100, text.size = 0.6)
# 
# 
# bigbatsHR <- 
#   tm_shape(bigbats_ranges, bbox = common_bbox) +
#   tm_borders(col = "salmon", lwd = 2, fill_alpha = 0.7) +
#   tm_fill(fill = "salmon", fill_alpha = 0.1) +
#   tm_shape(bigbats_points) +
#   tm_dots(fill = "salmon", size = 0.3, fill_alpha = 0.8) +
#   tm_basemap("Esri.WorldImagery") +
#   tm_scalebar(position = c("top", "right"), width = 200, text.size = 1.5)
# 
# tmap_mode("view")
# bigbatsHR
# 
# tmap_save(
#   bigbatsHR,
#   filename = file.path(output_today, "bigbatsHR_Sattelite.tiff"),
#   width = 8,
#   height = 12,
#   units = "in",
#   dpi = 600
# )
# 
# 
# 
#    tm_shape(Maren_sf, bbox = common_bbox) + 
#    tmap_mode('view') +
#    tm_basemap("CartoDB.Voyager") +
#    tm_dots(fill= "steelblue", size = .3, fill_alpha = .8) +
#  tm_shape(range95Maren_sf) + 
#    tm_borders(fill_alpha=.7, col = "steelblue", lwd = 2) + 
#    tm_fill(fill_alpha=.1, fill= "steelblue") 
#      
#    
#    tm_shape(Thea_sf, bbox = common_bbox) + 
#    tmap_mode('view') +
#    tm_basemap("CartoDB.Voyager") +
#    tm_dots(fill= "darkgreen", size = .3, fill_alpha = .8) +
#  tm_shape(range95Thea_sf) + 
#    tm_borders(fill_alpha=.7, col = "darkgreen", lwd = 2) + 
#    tm_fill(fill_alpha=.1, fill= "darkgreen") 
#    
#    tm_shape(Dagny_sf, bbox = common_bbox) + 
#    tmap_mode('view') +
#    tm_basemap("CartoDB.Voyager") +
#    tm_dots(fill= "tomato", size = .3, fill_alpha = .8) +
#    tm_shape(range95Dagny_sf) + 
#    tm_borders(fill_alpha=.7, col = "tomato", lwd = 2) + 
#    tm_fill(fill_alpha=.1, fill= "tomato")

```


