---
title: "2.RSF"
output: html_document
date: "2024-02-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
### Prepare the work space
# citation()
#install.packages(c("DHARMa", "lme4", "Matrix", "Rcpp", "RcppEigen"))

library(tidyverse)
library(dplyr)
library(scales)
library(beepr)
library(lubridate)
library(ResourceSelection)
library(RColorBrewer)
library(cowplot)
library(colorBlindness)
library(lme4)
library(stats)
library(DHARMa)
library(lattice)
library(broom)
library(modelsummary)
library(gt)
library(readxl)
library(xlsx)
library(corrplot)
library(vegan)
library(mapview)
library(sf)
library(influence.ME)


getwd()
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/Documents/1. PhD_Main/GitHub_link/Nittedal/SpatialAndMorphology_MBRA_MMYS"

output <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs"

file.name <- "RSF_50mRadius"

todays_date <- Sys.Date()
 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
dir.name
 
output_today <- dir.name
output_today

dir.create(output_today)
output_today
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs/RSF_50mRadius_2025-11-21"



```


```{r}
## Covariates from LiDAR data prepared by Hans Ole 
vars <- readRDS("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/SpatialData/231220_vars_dist50_v4_na-rm.rds")

#Used and available positions - just for referencing. 
rsf.data <- read_csv("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs/HomeRangeExploration_2023-09-28/used_available_pts_combined95KE.csv", 
      col_types = cols(...1 = col_skip()))

rsf.data$BatID <- factor(rsf.data$BatID)
used <- rsf.data %>% filter(used == 1) 

summary(used$BatID)

summary(vars)
str(vars) # sf data frame
# 16522 observations of 47 variables 

names(vars)
# [1] "BatID"             "Easting"           "Northing"          "used"              "Species"          
#  [6] "UTM"               "Hmax"              "Hmean"             "Hsd"               "Hcv"              
# [11] "Hkurt"             "Hskewness"         "H10"               "H20"               "H30"              
# [16] "H40"               "H50"               "H60"               "H70"               "H80"              
# [21] "H90"               "H95"               "D0"                "D1"                "D2"               
# [26] "D3"                "D4"                "D5"                "D6"                "D7"               
# [31] "D8"                "D9"                "gap3.n"            "gap3.mean.area"    "gap3.sd.area"     
# [36] "gap3.sum.area"     "gap5.n"            "gap5.mean.area"    "gap5.sd.area"      "gap5.sum.area"    
# [41] "ttops.n"           "crown.mean.size"   "crown.sd.size"     "ttops.mean.height" "ttops.sd.height"  
# [46] "water.distance"    "water.objtype"     "YOD"               "DIST"              "geom"           

## For writing descriptions of each variable 

#mettext <- as.data.frame(names(vars))
# write.csv(mettext, "~mettext.csv") # 15.11.2023

## First convert to a normal dataframe(drop geoms/sf objects)

dt <- vars
dt$BatID <- factor(dt$BatID) # 21 bats 
dt1df<-st_drop_geometry(dt)
dt1df$Species <- factor(dt1df$Species)
dt1df <- as.data.frame(dt)
dt1df <- dt1df %>% select(-geom)
# 16522 obs of 49 vars

levels(dt$BatID )

### FIX THIS HERE! Turid is somehow in the wrong category... also double check how the points are distributed between the two taxa!! 

## Inset bat UID and fix species error for Turid

dtx <- dt1df %>% mutate(
  UID = factor(case_when(
  BatID == "Amelia" ~ "M01",
  BatID == "Aricia" ~ "M11", 
  BatID == "Astrid" ~ "B07", 
  BatID == "Dagny" ~ "B09", 
  BatID == "Daisy" ~ "M06", 
  BatID == "Ethel" ~ "M08", 
  BatID == "Ida" ~ "M10", 
  BatID == "Kaja" ~ "B01", 
  BatID == "Line" ~ "M05", 
  BatID == "Louise" ~ "M12", 
  BatID == "Maren" ~ "B05", 
  BatID == "Marie" ~ "M02",
  BatID == "Nora" ~ "B03", 
  BatID == "Phoebe" ~ "B06", 
  BatID == "Ragnhild" ~ "M09", 
  BatID == "Reeda" ~ "M04", 
  BatID == "Sofia" ~ "B04", 
  BatID == "Steffi" ~ "B08", 
  BatID == "Stine" ~ "M07", 
  BatID == "Thea"  ~ "B02", 
  BatID == "Turid" ~ "M03")),
  species = factor(case_when(
    BatID == "Turid" ~ "M.mystacinus", 
    TRUE ~ Species
  ))) %>% select(-Species) %>% 
  mutate(Species = species) %>% 
  select(-species)

test <- dtx %>% select(BatID, Species, UID) %>% distinct() %>% arrange(UID)

test ## Fixed! 

summary(dtx)

table(dtx$BatID)
#   Amelia   Aricia   Astrid    Dagny    Daisy    Ethel      Ida 
#      176      990      891      913     1474     1287     1309 
#     Kaja     Line   Louise    Maren    Marie     Nora   Phoebe 
#     1001      605      704     1551      341      539      682 
# Ragnhild    Reeda    Sofia   Steffi    Stine     Thea    Turid 
#      308      770      429      429      924     1111       88

 
 ## Fixed! 


## There is an -Inf value for Hmax which can be corrrected like this: 
dtx$Hmax <- gsub("-Inf", "0.5", dtx$Hmax)
dtx$Hmax <- as.numeric(dtx$Hmax)
summary(dtx)

dt2df <- dtx 
dim(dt2df)
# 16522 obs of 50 vars

# Scale all continuous numerical data 
dt1 <- dt2df 
dt1[,6:45]=scale(dt1[,6:45], scale=TRUE) #dt)

dim(dt1)
# 16522 obs of 50 vars 

### Define non-forest areas in the dataset 
# How does this look for used points? 
# dt2dfused <- dt2df %>% filter(used == 1)
# hist(dt2dfused$ttops.n) 
# 
# hist(dt2dfused$D0)
# 
# table(dt1df$BatID)


# ## Define non-forest as those with H95 < 5 and D0 less than 10 for used locations
# openareas <- dt2dfused %>% dplyr::filter(H95 < 5 & D0 < 0.1)
# summary(openareas)
# dim(openareas)
# # 22 obs 
# 22/1502
# # 0.014
# 0.014 *100
# 
# table(openareas$Species)
#   # M.brandtii M.mystacinus 
#   #         15            7 
# 
# table(openareas$BatID)
# #   Amelia   Aricia   Astrid    Dagny    Daisy    Ethel      Ida 
# #        2        0        0        1        0        2        0 
# #     Kaja     Line   Louise    Maren    Marie     Nora   Phoebe 
# #        0        0        0        1        0        0        5 
# # Ragnhild    Reeda    Sofia   Steffi    Stine     Thea    Turid 
# #        0        1        0        7        2        1        0
# 
# ggplot(dt2dfused, aes(x = H95, y = D0)) + 
#   geom_point(color = "orange", alpha = 0.7) + geom_smooth(color = "black") +
#   facet_wrap(~Species)
# 
# ## WHat is the proportion of non forest areas for random pointS? 
# random <- dt2df %>% filter(used == 0)
# 
# ## Define non-forest as those with H95 < 5 and D0 less than 10 for used locations
# openareas <- random %>% dplyr::filter(H95 < 5 & D0 < 0.1)
# summary(openareas)
# dim(openareas)
# # 1677 obs 
# 1677/15020
# # 11%
# 
# table(openareas$Species)
#   # M.brandtii        M.mystacinus 
#   #          1016             661 
# 
# table(openareas$BatID)
# #   Amelia   Aricia   Astrid    Dagny    Daisy    Ethel      Ida 
# #       29      101       87      173       84      111        8 
# #     Kaja     Line   Louise    Maren    Marie     Nora   Phoebe 
# #       78       36       77      243       32       54       84 
# # Ragnhild    Reeda    Sofia   Steffi    Stine     Thea    Turid 
# #       26       10       27       67      147      203        0 
# 
# ggplot(random, aes(x = H95, y = D0)) + 
#   geom_point(color = "orange", alpha = 0.7) + geom_smooth(color = "black") +
#   facet_wrap(~Species)
 
```


## Correlation/collinearity between LiDAR variables 
```{r}
# names(dt1)
# dtc <- dt1 %>% dplyr::select(c(
#   Hmax, Hmean, Hsd, Hcv, Hskewness, 
#   H10, H20, H30, H40,  H50, H60, H70, H80, H90, H95, 
#   D0, D1, D2, D3, D4, D5, D6, D7, D8, D9, 
#   gap3.n, gap3.mean.area,  gap3.sd.area,  gap3.sum.area,  
#   gap5.n, gap5.mean.area,  gap5.sd.area,  gap5.sum.area,  
#   ttops.n, ttops.mean.height, ttops.sd.height)) 
# 
# windows()
# 
# corrplot(cor(dtc))
```


# PCA ordination of available points to address collinearity and discern patterns in how variables influence each other 


### Prepare data for PCA 
```{r}

## Create an ID column so the bat meta data can be reattached later

# ## Create a unique row id which will become "idvar"
# dt1$idvar <- paste0("ID-", 1:nrow(dt1)) 
# headings <- dt1$idvar
# 
# ## Now only includeLiDAR variables with IDvar
# names(dt1)
# 
# dt2 <- dt1 %>% dplyr::select(c(
#   Hmax, Hmean, Hsd, Hcv, Hskewness, Hkurt, 
#   H10, H20, H30, H40,  H50, H60, H70, H80, H90, H95, 
#   D0, D1, D2, D3, D4, D5, D6, D7, D8, D9, 
#   gap3.n, gap3.mean.area,  gap3.sd.area,  gap3.sum.area,  
#   gap5.n, gap5.mean.area,  gap5.sd.area,  gap5.sum.area,  
#   ttops.n, ttops.mean.height, ttops.sd.height,
#   crown.mean.size, crown.sd.size)) 
# 
# rownames(dt2) <- headings
# # Reshape into wide format 
# 
# summary(dt2)
# ## This is a PCA, more info below: 
# 
# mat <- dt2 

```

## PCA
```{r}

# ## PCA 
# trait.pca <- rda(mat) # used and available points together 
# biplot(trait.pca) 
# # simple plot 
# screeplot(trait.pca) 
# # how much of each axis describes the variation in the data 
# # PC1 Explains most of the variation. Big difference between PC2 and PC3. 
# summary(trait.pca)
# pca.summary <- summary(trait.pca)
# ## In our case, 'species' are the different LiDAR variables and sites are the different bat locations. 
# trait.loadings <- as.data.frame(pca.summary$species) %>% 
#   mutate(trait = rownames(.))  
# head(trait.loadings)
# meta <- dt1 %>% dplyr::select(c(used, Species, idvar)) 
# # for now, do not include bat.id
# ## Build meta data to be reattached in plotting
# rownames(meta) <- meta$idvar
# meta$gen.spe <- meta$idvar
# meta$used <- as.factor(meta$used)
# # Scores for all the different bat locations 
# 
# sp.scores <- as.data.frame(pca.summary$sites) %>% 
#   dplyr::mutate(gen.spe = rownames(.)) %>% dplyr::inner_join(meta, by="gen.spe") 
# 
# traittable <- trait.loadings  %>% 
#   dplyr::select(trait, PC1, PC2)
# 
# #write.csv(traittable, file = file.path(output_today, "bothbats_allpts_PCA1_2_table.csv")) # 19.11.2025
# 
# ### Compare trait loadings for bat locations with Pearson's correlation of PC1 and PC2, as well as of the LiDAR variables with each other 
# spdf <- dplyr::left_join(sp.scores, dt1, by = "idvar")
# spdf1 <- spdf %>% dplyr::select(-c(PC3, PC4, PC5, PC6, idvar, gen.spe, used.x, used.y, Species.x, Species.y, BatID, UID, Easting, Northing, UTM, YOD, DIST, water.distance, water.objtype))
# #corrplot(cor(dtc))
# cortab <- as.data.frame(cor(spdf1)) 
# #windows()
# corrplot(cor(spdf1),
#          tl.col="black", tl.srt=45,  
#          type="lower", col=brewer.pal(n=5, name="Greys"))
# cortab1 <- cortab[1:2,]
# cortab2 <- t(cortab1) # Table of the values for further comparing 
# # write.csv(cortab2, file = file.path(output_today, "cortab2.csv"))
# # 21.11.2025
# 
# ################################################################################
# ##### plotting 
# ################################################################################
# 
# ## PC1 and PC2
# pca.ggplot <- ggplot(data = sp.scores, aes(x = PC1, y = PC2)) +
#   geom_point() +
#   geom_segment(data = trait.loadings, 
#                aes(x = 0, xend = PC1/2, y = 0, yend = PC2/2), lwd = 0.1) + # this is a new geom that allows you to draw segments like the ones for the loadings above
#   geom_text(data = trait.loadings, aes(x = PC1/2, y = PC2/2+0.1, label = trait), size = 3) + # this adds text to the labels
#   theme_bw() 
# pca.ggplot
# 
# 
# ## Table of PCA for further informing variable selection 
# traittable <- trait.loadings  %>% 
#   dplyr::select(trait, PC1, PC2)
# # write.csv(traittable, file = file.path(output_today, "bothbats_allpts_PCA1_2_table.csv")) # 24.11.2025
# 
# ## Now create a similar table for MBRA and MMYS separately 


```


## RSF
```{r}
## Simple models - separated by bat species, no species interaction 
# install.packages("lme4", type = "source") 
# library(lme4)

## 19.11.2025
# I discovered that Turid has erroneously been placed with M. brandtii when in fact she should be with M. mystacinus 
## So now I am going to have to re-define the species column 

## start here 
# fix <- dt1 %>% mutate(species = factor(
#   case_when(
#     BatID == "Turid" ~ "M.mystacinus"),
#     FALSE ~ Species)) 

MBRA <- dt1 %>% filter(Species == "M.brandtii") %>% droplevels() # 7634 obs of 49 vars
MMYS <- dt1 %>% filter(Species == "M.mystacinus") %>% droplevels() # 8888 obs of 49 vars

MBRA$BatID <- factor(MBRA$BatID) # 9 levels 
table(MBRA$BatID)
# Astrid  Dagny   Kaja  Maren   Nora Phoebe  Sofia Steffi   Thea 
#    891    913   1001   1551    539    682    429    429   1111 


MMYS$BatID <- factor(MMYS$BatID) # 11 levels 
table(MMYS$BatID)
  # Amelia   Aricia    Daisy    Ethel      Ida     Line   Louise 
  #    176      990     1474     1287     1309      605      704 
  #  Marie Ragnhild    Reeda    Stine    Turid 
  #    341      308      770      924       88 

summary(MMYS)
summary(MBRA)

########################################################
########################  MBRA  ########################
########################################################
# install.packages(c("Matrix", "Rcpp", "RcppEigen", "lme4"))
library(lme4)
fit1MBRA= glmer(used ~
                  gap3.mean.area +
                  D0 + H95 + Hsd + water.distance +
                  (1|BatID), data=MBRA,
              family=binomial(link="logit"),nAGQ = 0)
fit2MBRA= glmer(used ~
                  gap5.mean.area +
                  D0 + H95 + Hsd + water.distance +
                  (1|BatID), data=MBRA,
              family=binomial(link="logit"),nAGQ = 0)
# 
summary(fit2MBRA)

isSingular(fit1MBRA)
# # FALSE
# isSingular(fit2MBRA)
# # FALSE
# AIC(fit1MBRA, fit2MBRA)
#          df      AIC
# fit1MBRA  7 4290.572 # 3 m scale is lower 
# fit2MBRA  7 4301.640

# (old analysis results)
# fit1MBRA  7 4348.591 # 3 m scale is lower 
# fit2MBRA  7 4359.372 

# fit 2 has a lower AIC weight BUT it gets flagged for having a singular fit so I would instead want to use the model with the 3 m resolution gap variables. 
########################################################
########################  MMYS  ########################
########################################################
# 
# 
fit1MMYS= glmer(used ~
                  gap3.mean.area +
                  D0 + H95 + Hsd + water.distance +
                  (1|BatID), data=MMYS,
              family=binomial(link="logit"),nAGQ = 0)
fit2MMYS= glmer(used ~
                  gap5.mean.area +
                  D0 + H95 + Hsd + water.distance +
                  (1|BatID), data=MMYS,
              family=binomial(link="logit"),nAGQ = 0)
AIC(fit1MMYS, fit2MMYS)
#          df      AIC
# fit1MMYS  7   5201.882
# fit2MMYS  7   5199.381 # 5 m scale is lower


## (old analysis)
# fit1MMYS  7 5145.626
# fit2MMYS  7 5143.210 # 5 m scale is lower 
         # The difference between the weights is negligible, I feel comfortable sticking with the 3 m resolution due to the singularity issues found in the M. brandtii models. 
################################################
# The gap5 model for MBRA throws a singularity warning, and the AIC weights between the model pairs is very little so I will use the 3 m resolution models to continue. 

# summary(fit1MBRA)
# 
# summary(fit1MMYS)
# 
sjPlot::tab_model(fit1MBRA, fit1MMYS,fit2MBRA, fit2MMYS,
                  dv.labels = c(
                    "Myotis brandtii - 3 m", "Myotis mystacinus - 3 m",
                    "Myotis brandtii - 5 m", "Myotis mystacinus - 5 m"), transform = NULL, auto.label = FALSE)

# Plot the confidence intervals for MBRA and MMYS at 5 m resolution


```


## Use the VIF function to test for correlation amongst variables 
```{r}
# library(car)
### VIF equal to 1 = variables are not correlated.
## VIF between 1 and 5 = variables are moderately correlated.
## VIF greater than 5 = variables are highly correlated.


# vif(fit1MBRA)
# gap3.mean.area             D0            H95            Hsd water.distance 
#       2.728157       2.409865       6.196688       5.262636       1.154840 

## Old analysis: 
# gap3.mean.area             D0            H95            Hsd water.distance 
#       2.741990       2.413678       6.131832       5.201883       1.152878 

# vif(fit1MMYS)
# gap3.mean.area             D0            H95            Hsd water.distance 
#       2.464794       2.996740       4.653837       3.828337       1.233227 

## Old analysis: 
# gap3.mean.area             D0            H95            Hsd water.distance 
#       2.464688       3.025031       4.715925       3.863640       1.249359 
# (bigfit2)

# Fixed effects:
#                                    Estimate Std. Error z value Pr(>|z|)    
# (Intercept)                        -2.36989    0.06354 -37.300  < 2e-16 ***
# SpeciesM.mystacinus                -0.20330    0.08833  -2.302   0.0214 *  
# gap3.mean.area                     -0.61737    0.07116  -8.676  < 2e-16 ***
# D0                                 -0.44494    0.07080  -6.284 3.30e-10 ***
# Hsd                                 0.25406    0.04561   5.570 2.54e-08 ***
# water.distance                     -0.51920    0.03588 -14.471  < 2e-16 ***
# SpeciesM.mystacinus:gap3.mean.area  0.43080    0.09966   4.323 1.54e-05 ***
# SpeciesM.mystacinus:D0              0.60367    0.09596   6.291 3.16e-10 ***
# SpeciesM.mystacinus:Hsd             0.13839    0.06488   2.133   0.0329 *  
   
# Drop H95 to see how this changes the correlation amongst variables 
# fit3MBRA= glmer(used ~  
#                   gap3.mean.area + 
#                   D0 + Hsd + water.distance +
#                   (1|BatID), data=MBRA,
#               family=binomial(link="logit"),nAGQ = 0) 
# fit3MMYS= glmer(used ~  
#                   gap3.mean.area + 
#                   D0 + Hsd + water.distance +
#                   (1|BatID), data=MMYS,
#               family=binomial(link="logit"),nAGQ = 0) 
# 
# 
# vif(fit3MBRA)
# gap3.mean.area             D0            Hsd water.distance 
#       2.500579       2.330923       1.117494       1.135345 

## Old analysis: 
# gap3.mean.area             D0            Hsd water.distance 
#       2.513972       2.334639       1.111466       1.135041  


# vif(fit3MMYS)
# gap3.mean.area             D0            Hsd water.distance 
#       2.500702       2.590675       1.074616       1.224396 

## Old analysis: 
# gap3.mean.area             D0            Hsd water.distance 
#       2.503120       2.605929       1.076558       1.235351 

# test <- dt1 %>% select(BatID, Species, UID) %>% distinct() %>% arrange(UID)
# test

## Sorted 
# bigfit1 = glmer(used ~  Species + 
#                   gap3.mean.area + 
#                   D0 + Hsd + water.distance +
#                   Species*gap3.mean.area + 
#                   Species*D0 + 
#                   Species*Hsd + 
#                   Species*water.distance +
#                   (1|BatID), data=dt1,
#               family=binomial(link="logit"),nAGQ = 0)
# summary(bigfit1)
# ## Now step-wise selection with AIC 
# drop1(bigfit1)
# Model:
# used ~ Species + gap3.mean.area + D0 + Hsd + water.distance + 
#     Species * gap3.mean.area + Species * D0 + Species * Hsd + 
#     Species * water.distance + (1 | BatID)
#                        npar    AIC
# <none>                      9565.0
# Species:gap3.mean.area    1 9581.2
# Species:D0                1 9604.9
# Species:Hsd               1 9567.4
# Species:water.distance    1 9563.3 ## this is low enough still that it warrants dropping 

## Old analysis:
# used ~ Species + gap3.mean.area + gap3.sd.area + D0 + Hsd + water.distance + 
#     Species * gap3.mean.area + Species * gap3.sd.area + Species * 
#     D0 + Species * Hsd + Species * water.distance + (1 | BatID)
# 
#                        npar    AIC
# <none>                      9565.6
# Species:gap3.mean.area    1 9581.0
# Species:D0                1 9604.8
# Species:Hsd               1 9567.4
# Species:water.distance    1 9564.5 ** 

# Drop interaction between species and distance to water
bigfit2 =  glmer(used ~  Species + 
                   gap3.mean.area + 
                   D0 + Hsd + water.distance +
                   Species*gap3.mean.area + 
                   Species*D0 + 
                   Species*Hsd + 
                   (1|BatID), data=dt1,
               family=binomial(link="logit"),nAGQ = 0)

# 
# drop1(bigfit2)
# boundary (singular) fit: see help('isSingular')
# used ~ Species + gap3.mean.area + D0 + Hsd + water.distance + 
#     Species * gap3.mean.area + Species * D0 + Species * Hsd + 
#     (1 | BatID)
## Suggests to keep everything 
#                        npar    AIC
# <none>                      9563.3
# water.distance            1 9786.3
# Species:gap3.mean.area    1 9580.2
# Species:D0                1 9603.0
# Species:Hsd               1 9566.2

summary(bigfit2)
# Generalized linear mixed model fit by maximum likelihood (Adaptive Gauss-Hermite
#   Quadrature, nAGQ = 0) [glmerMod]
#  Family: binomial  ( logit )
# Formula: used ~ Species + gap3.mean.area + D0 + Hsd + water.distance +  
#     Species * gap3.mean.area + Species * D0 + Species * Hsd +      (1 | BatID)
#    Data: dt1
# 
#      AIC      BIC   logLik deviance df.resid 
#   9563.3   9640.4  -4771.6   9543.3    16512 
# 
# Scaled residuals: 
#     Min      1Q  Median      3Q     Max 
# -0.7635 -0.3560 -0.2783 -0.2008  7.8062 
# 
# Random effects:
#  Groups Name        Variance Std.Dev.
#  BatID  (Intercept) 0.0187   0.1368  
# Number of obs: 16522, groups:  BatID, 21
# 
# Fixed effects:
#                                    Estimate Std. Error z value Pr(>|z|)    
# (Intercept)                        -2.37265    0.06453 -36.768  < 2e-16 ***
# SpeciesM.mystacinus                -0.19439    0.08878  -2.189   0.0286 *  
# gap3.mean.area                     -0.62317    0.07155  -8.709  < 2e-16 ***
# D0                                 -0.45490    0.07149  -6.364 1.97e-10 ***
# Hsd                                 0.25243    0.04578   5.514 3.51e-08 ***
# water.distance                     -0.51682    0.03585 -14.418  < 2e-16 ***
# SpeciesM.mystacinus:gap3.mean.area  0.43776    0.09973   4.390 1.14e-05 ***
# SpeciesM.mystacinus:D0              0.61467    0.09615   6.393 1.63e-10 ***
# SpeciesM.mystacinus:Hsd             0.14392    0.06491   2.217   0.0266 *  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Correlation of Fixed Effects:
#             (Intr) SpcsM. gp3.m. D0     Hsd    wtr.ds SM.:3. SM.:D0
# SpcsM.mystc -0.709                                                 
# gap3.mean.r  0.030 -0.003                                          
# D0           0.100 -0.076  0.737                                   
# Hsd         -0.157  0.138  0.231  0.074                            
# water.dstnc  0.126  0.054  0.127 -0.022  0.167                     
# SpcsM.m:3.. -0.019  0.084 -0.715 -0.529 -0.162 -0.070              
# SpcsM.my:D0 -0.090  0.070 -0.563 -0.741 -0.075 -0.104  0.757       
# SpcsM.mys:H  0.119 -0.204 -0.155 -0.054 -0.695 -0.053  0.227  0.122

sjPlot::tab_model(bigfit2, transform = NULL, auto.label = FALSE, show.re.var = TRUE,   show.fstat = TRUE, show.stat = TRUE) 

```



```{r}
### Model diagnostics for 
simulationOutput <- simulateResiduals(fittedModel = bigfit2, plot = T)
### Help with interpretation of residuals vs. predicted: 
# https://stats.stackexchange.com/questions/449208/diagnostic-plot-residual-vs-predicted-of-a-glmm-using-dharma
### recalculateResiduals()
# KS test -  the p-value shows you that there is *not* a significant deviation from the assumed distribution
# Dispersion test – Not over or underdispersed
# Outlier – No indication of outliers 
# 
# Residuals appear to more or less agree with predicted values 
#windows()
testDispersion(simulationOutput)
# 
# 	DHARMa nonparametric dispersion test via sd of residuals fitted vs.
# 	simulated
# 
# data:  simulationOutput
# dispersion = 0.9964, p-value = 0.944
# alternative hypothesis: two.sided
# no strong evidence for Dispersion issues 
testZeroInflation(simulationOutput)
# 	DHARMa zero-inflation test via comparison to expected zeros with simulation
# 	under H0 = fitted model
# 
# data:  simulationOutput
# ratioObsSim = 1.0012, p-value = 0.816
# alternative hypothesis: two.sided
# No strong evidence for zero inflation 
testQuantiles(simulationOutput)
# 	Test for location of quantiles via qgam
# 
# 	Test for location of quantiles via qgam
# 
# data:  simulationOutput
# p-value = 0.147
# alternative hypothesis: both

testCategorical(simulationOutput, catPred = dt1$BatID)
## No bat IDs had a low p value for the Asymptotic  one-sample Kolmogorov-Smirnov test
testCategorical(simulationOutput, catPred = dt1$Species)
## No bat species had a low p value for the Asymptotic  one-sample Kolmogorov-Smirnov test

vif(bigfit2)
            #    Species         gap3.mean.area                     D0 
            #   1.093962               5.201610               5.476647 
            #        Hsd         water.distance Species:gap3.mean.area 
            #   2.125315               1.171877               5.317849 
            # Species:D0            Species:Hsd 
            #   5.645544               2.141335 

### Old analysis: 
        # Species         gap3.mean.area                     D0                    
#       1.095477               5.148412               5.380427    
#       Species:D0       water.distance          Species:gap3.mean.area             
#        2.109284              1.174590               5.280176                       
#     Species:D0            Species:Hsd 
#      5.574119                2.127375 


# Higher VIF values for interacting terms is not too much of a concern 

```


## Plot predictions using glm 
```{r}
## Refitting the model with GLM to sketch results
bigfit2.0<- glm(used ~  Species + 
                   gap3.mean.area + 
                   D0 + Hsd + water.distance +
                   Species*gap3.mean.area + 
                   Species*D0 + 
                   Species*Hsd, 
                  family=binomial(), data=dt1)
summary(bigfit2.0)

# Call:
# glm(formula = used ~ Species + gap3.mean.area + D0 + Hsd + water.distance + 
#     Species * gap3.mean.area + Species * D0 + Species * Hsd, 
#     family = binomial(), data = dt1)
# 
# Coefficients:
#                                    Estimate Std. Error z value Pr(>|z|)    
# (Intercept)                        -2.35578    0.04401 -53.527  < 2e-16 ***
# SpeciesM.mystacinus                -0.20635    0.06151  -3.355 0.000794 ***
# gap3.mean.area                     -0.65471    0.06869  -9.531  < 2e-16 ***
# D0                                 -0.46646    0.07124  -6.547 5.85e-11 ***
# Hsd                                 0.24381    0.04473   5.451 5.00e-08 ***
# water.distance                     -0.48438    0.03488 -13.888  < 2e-16 ***
# SpeciesM.mystacinus:gap3.mean.area  0.48231    0.09741   4.952 7.36e-07 ***
# SpeciesM.mystacinus:D0              0.59449    0.09545   6.228 4.72e-10 ***
# SpeciesM.mystacinus:Hsd             0.14181    0.06383   2.222 0.026297 *  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# (Dispersion parameter for binomial family taken to be 1)
# 
#     Null deviance: 10066.4  on 16521  degrees of freedom
# Residual deviance:  9550.5  on 16513  degrees of freedom
# AIC: 9568.5
# 
# Number of Fisher Scoring iterations: 5


# Number of Fisher Scoring iterations: 6
batcolors = c("#32A5FF", "#011f4b") 

# New facet label names for supp variable
supp.labs <- c("M. brandtii", "M. mystacinus")
names(supp.labs) <- c("M.brandtii", "M.mystacinus")
### Plotting predictions ####
# gap3.mean ---------------------------------------
range.gm <- range(dt1$gap3.mean.area)
range.gm
# -1.040815  1.198892
  
pred <- data.frame(expand.grid(
  D0 = mean(dt1$D0),
  gap3.mean.area = seq(from = -1.040815  , to = 1.198892, by = 0.1),
  Hsd = mean(dt1$Hsd), 
  water.distance = mean(dt1$water.distance),
  Species = c("M.brandtii", "M.mystacinus"))) 
  
pred <- cbind(pred, predict(bigfit2.0, newdata = pred,
                             type = "link", se.fit = TRUE))
pred$PP <- plogis(pred$fit)
pred$LL <- plogis(pred$fit - (1.96 * pred$se.fit))
pred$UL <- plogis(pred$fit + (1.96 * pred$se.fit))
## Plot 
p1 <- ggplot(pred, aes(gap3.mean.area, PP)) + 
  geom_line(aes(col = factor(Species)), linewidth = 1.5) + 
  geom_ribbon(aes(x = gap3.mean.area, ymin = LL, ymax = UL,
              fill = factor(Species)), alpha = 0.4)  +
  facet_wrap(~Species, labeller = labeller(Species = supp.labs)) +
  scale_y_continuous(
  labels = scales::number_format(accuracy = 0.01,
                                 decimal.mark = '.')) +
  theme_bw() +
  theme(strip.text = element_text(face = "italic"),
        legend.position="none") +
  scale_colour_manual(values = 
                        batcolors,  
                      aesthetics = c("fill", "colour"), 
                      name = "Bat species") + 
  xlab ("Mean gap area") +
  ylab ("") +
  theme(text = element_text(size=20))
p1

# distance to water ---------------------------------------
range.gs <- range(dt1$water.distance)
range.gs
#-1.282820  3.698758
  
pred <- data.frame(expand.grid(
  D0 = mean(dt1$D0),
  gap3.mean.area = mean(dt1$gap3.mean.area),
  Hsd = mean(dt1$Hsd), 
  water.distance = seq(from = -1.282820  , to = 3.698758, by = 0.1),
  Species = c("M.brandtii", "M.mystacinus"))) 
  
pred <- cbind(pred, predict(bigfit2.0, newdata = pred,
                             type = "link", se.fit = TRUE))
pred$PP <- plogis(pred$fit)
pred$LL <- plogis(pred$fit - (1.96 * pred$se.fit))
pred$UL <- plogis(pred$fit + (1.96 * pred$se.fit))

## Plot 
p2 <- ggplot(pred, aes(water.distance, PP)) + 
  geom_line(aes(col = factor(Species)), linewidth = 1.5) + 
  geom_ribbon(aes(x = water.distance, ymin = LL, ymax = UL, 
              fill = factor(Species)), alpha = 0.4) +
  facet_wrap(~Species,labeller = labeller(Species = supp.labs)) +
  scale_y_continuous(
  labels = scales::number_format(accuracy = 0.01,
                                 decimal.mark = '.')) +
  theme_bw() +
  theme(strip.text = element_text(face = "italic"),
        legend.position="none") +
  scale_colour_manual(values = 
                        batcolors,  
                      aesthetics = c("fill", "colour"), 
                      name = "Bat species") + 
  xlab ("Distance to water") +
  ylab ("") +
  theme(text = element_text(size=20))
p2


# D0---------------------------------------
range.D0 <- range(dt1$D0)
range.D0
# -0.9838802  3.3390781
  
pred <- data.frame(expand.grid(
  D0 = seq(from = -0.9838802, to = 3.3390781, by = 0.1),
  gap3.sd.area = mean(dt1$gap3.sd.area),
  gap3.mean.area = mean(dt1$gap3.mean.area),
  Hsd = mean(dt1$Hsd), 
  water.distance = mean(dt1$water.distance),
  Species = c("M.brandtii", "M.mystacinus"))) 
  
pred <- cbind(pred, predict(bigfit2.0, newdata = pred,
                             type = "link", se.fit = TRUE))
pred$PP <- plogis(pred$fit)
pred$LL <- plogis(pred$fit - (1.96 * pred$se.fit))
pred$UL <- plogis(pred$fit + (1.96 * pred$se.fit))
## Plot 
p3 <- ggplot(pred, aes(D0, PP)) + 
  geom_line(aes(col = factor(Species)), linewidth = 1.5) + 
  geom_ribbon(aes(x = D0, ymin = LL, ymax = UL,
                  fill = factor(Species)), alpha = 0.4) + 
  facet_wrap(~Species, labeller = labeller(Species = supp.labs)) +
  scale_y_continuous(
  labels = scales::number_format(accuracy = 0.01,
                                 decimal.mark = '.')) +
  theme_bw() +
  theme(strip.text = element_text(face = "italic"),
        legend.position="none") +
  scale_colour_manual(values = 
                        batcolors,  
                      aesthetics = c("fill", "colour"), 
                      name = "Bat species") + 
  ylab ("") +
  xlab ("D0") +
  theme(text = element_text(size=20))
p3

# Hsd ---------------------------------------
range.Hsd <- range(dt1$Hsd)
range.Hsd

#Now: 
# -2.233849  3.866324

# Previously 
# -2.125363  3.625778
  
pred <- data.frame(expand.grid(
  D0 = mean(dt1$D0),
  gap3.sd.area = mean(dt1$gap3.sd.area),
  gap3.mean.area = mean(dt1$gap3.mean.area),
  Hsd = seq(from = -2.233849, to = 3.866324, by = 0.1), 
  water.distance = mean(dt1$water.distance),
  Species = c("M.brandtii", "M.mystacinus"))) 
  
pred <- cbind(pred, predict(bigfit2.0, newdata = pred,
                             type = "link", se.fit = TRUE))
pred$PP <- plogis(pred$fit)
pred$LL <- plogis(pred$fit - (1.96 * pred$se.fit))
pred$UL <- plogis(pred$fit + (1.96 * pred$se.fit))
## Plot 
p4 <- ggplot(pred, aes(Hsd, PP)) + 
  geom_line(aes(col = factor(Species)), linewidth = 1.5) + 
  geom_ribbon(aes(x = Hsd, ymin = LL, ymax = UL, 
                  fill = factor(Species)), alpha = 0.4) + 
  facet_wrap(~Species, labeller = labeller(Species = supp.labs)) +
  scale_y_continuous(
  labels = scales::number_format(accuracy = 0.01,
                                 decimal.mark = '.')) +
  theme_bw() +
  theme(strip.text = element_text(face = "italic"),
        legend.position="none") +
  scale_colour_manual(values = 
                        batcolors,  
                      aesthetics = c("fill", "colour"), 
                      name = "Bat species") + 
  xlab ("Hsd") +
  ylab("") +
  theme(text = element_text(size=20))
p4

library(cowplot)
windows()
cowplot::plot_grid(p1, p2, p3, p4, nrow = 4, 
                   labels = c("a.", "b.", "c.", "d."), 
                   label_x = 0.08, label_y = 0.08,
                   hjust = -0.5, vjust = -0.5) +
  theme(text = element_text(size=16)) 

ggsave(path = output_today, filename = "Fig3.tiff", width = 9, height = 12, device='tiff', dpi=600)


cowplot::plot_grid(p1, p2, p3, p4, nrow = 2, 
                   labels = c("a.", "b.", "c.", "d."), 
                   label_x = 0.08, label_y = 0.08,
                   hjust = -0.5, vjust = -0.5) +
  theme(text = element_text(size=16)) 

ggsave(path = output_today, filename = "Fig3_forKappa.tiff", width = 10, height = 7, device='tiff', dpi=300)


output_today
```



## Further troubleshoot the model to test the robustness
How much does a single individual influence the outcome of the model? 


bigfit2 =  glmer(used ~  Species + 
                   gap3.mean.area + 
                   D0 + Hsd + water.distance +
                   Species*gap3.mean.area + 
                   Species*D0 + 
                   Species*Hsd + 
                   (1|BatID), data=dt1,
               family=binomial(link="logit"),nAGQ = 0)

```{r}


# =============================================================================
# 2. COOK'S DISTANCE AND DFBETAS (influence.ME)
# =============================================================================
# This refits the model n times, dropping one individual per iteration.
# Can be slow for large datasets; set parallel = TRUE if using a cluster.

infl <- influence(bigfit2, group = "BatID")
cd <- cooks.distance(infl)

cd_df <- data.frame(
  BatID = rownames(cd),
  cooks_distance = as.numeric(cd)
)

cd_df

#       BatID cooks_distance
# 1    Amelia    0.087254669
# 2    Aricia    0.466942487
# 3    Astrid    0.278351495
# 4     Dagny    0.645118904
# 5     Daisy    0.302865800
# 6     Ethel    0.644116758
# 7       Ida    0.329554532
# 8      Kaja    0.116871787
# 9      Line    0.210178012
# 10   Louise    0.518241172
# 11    Maren    1.476387292
# 12    Marie    0.173244284
# 13     Nora    0.156122941
# 14   Phoebe    0.179989711
# 15 Ragnhild    0.121511521
# 16    Reeda    0.311126351
# 17    Sofia    0.213415269
# 18   Steffi    0.151305957
# 19    Stine    0.532828146
# 20     Thea    4.550412472
# 21    Turid    0.007731942

               
               
        summary(dt1)

# --- Cook's distance ---
cook_d <- cooks.distance(infl)
cook_df <- data.frame(
  individual_id = rownames(cook_d),
  cooks_d       = as.numeric(cook_d),
  species       = dt1 |>
    distinct(BatID, Species) |>
    arrange(BatID) |>
    pull(Species)
)

cook_df


# Threshold: common rule of thumb is 4/n (conservative) or visual inspection
threshold_cook <- 4 / 21

p_cook <- ggplot(cook_df, aes(x = reorder(individual_id, cooks_d),
                               y = cooks_d, fill = species)) +
  geom_col() +
  geom_hline(yintercept = threshold_cook, linetype = "dashed", colour = "red") +
  coord_flip() +
  labs(
    title = "Cook's distance by individual",
    subtitle = paste0("Dashed line = 4/n threshold (", round(threshold_cook, 3), ")"),
    x = NULL, y = "Cook's D"
  ) +
  theme_bw()

print(p_cook)
ggsave("cooks_distance.png", p_cook, width = 7, height = 5, dpi = 300)


temp <- dt1 %>% mutate(individual_id = BatID, 
                       species = Species) 
# --- DFBETAS ---
dfb <- dfbetas(infl)
dfb_df <- as.data.frame(dfb) |>
  mutate(individual_id = rownames(dfb)) |>
  pivot_longer(-individual_id, names_to = "coefficient", values_to = "dfbeta") |>
  left_join(distinct(temp, individual_id, species), by = "individual_id")

# DFBETAS threshold: |DFBETAS| > 2/sqrt(n)
threshold_dfb <- 2 / sqrt(21)

p_dfb <- ggplot(dfb_df, aes(x = reorder(individual_id, abs(dfbeta)),
                              y = dfbeta, colour = species)) +
  geom_point(size = 2) +
  geom_hline(yintercept =  threshold_dfb, linetype = "dashed", colour = "red") +
  geom_hline(yintercept = -threshold_dfb, linetype = "dashed", colour = "red") +
  facet_wrap(~ coefficient, scales = "free_y") +
  coord_flip() +
  labs(
    title = "DFBETAS by individual and coefficient",
    subtitle = paste0("Dashed lines = ±2/√n threshold (±", round(threshold_dfb, 3), ")"),
    x = NULL, y = "DFBETAS"
  ) +
  theme_bw(base_size = 9)

print(p_dfb)
ggsave("dfbetas.png", p_dfb, width = 10, height = 7, dpi = 300)

```



```{r}
# =============================================================================
# 3. LEAVE-ONE-OUT (LOO) JACKKNIFE BY INDIVIDUAL
# =============================================================================
# Refit model dropping one individual at a time; record fixed-effect estimates.

# bigfit2 =  glmer(used ~  Species + 
#                    gap3.mean.area + 
#                    D0 + Hsd + water.distance +
#                    Species*gap3.mean.area + 
#                    Species*D0 + 
#                    Species*Hsd + 
#                    (1|BatID), data=dt1,
#                family=binomial(link="logit"),nAGQ = 0)


str(temp)
dat <- temp

individuals <- unique(dat$individual_id)

loo_results <- lapply(individuals, function(id) {
  dat_sub <- filter(dat, individual_id != id)
  m_loo <- tryCatch(
    glmer(
      used ~ Species + gap3.mean.area + 
                    D0 + Hsd + water.distance +
                    Species*gap3.mean.area + 
                    Species*D0 + 
                    Species*Hsd + 
             (1 | individual_id),
      data   = dat_sub,
      family = binomial(link = "logit"),
      control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
    ),
    error = function(e) NULL
  )
  if (is.null(m_loo)) return(NULL)
  coef_df <- as.data.frame(t(fixef(m_loo)))
  coef_df$dropped_individual <- id
  coef_df$dropped_species    <- dat |>
    filter(individual_id == id) |>
    pull(species) |>
    unique()
  coef_df
}) |>
  bind_rows()

# Full-model estimates for reference line
full_coefs <- data.frame(t(fixef(bigfit2)))
full_coefs$dropped_individual <- "FULL MODEL"
full_coefs$dropped_species    <- NA

# Combine and reshape
loo_long <- bind_rows(loo_results, full_coefs) |>
  pivot_longer(
    cols      = -c(dropped_individual, dropped_species),
    names_to  = "coefficient",
    values_to = "estimate"
  )

# Plot: one panel per coefficient, each point is the LOO estimate
p_loo <- loo_long |>
  filter(dropped_individual != "FULL MODEL") |>
  ggplot(aes(x = reorder(dropped_individual, estimate),
             y = estimate, colour = dropped_species)) +
  geom_point(size = 2) +
  geom_hline(
    data = filter(loo_long, dropped_individual == "FULL MODEL"),
    aes(yintercept = estimate),
    linetype = "dashed", colour = "black"
  ) +
  facet_wrap(~ coefficient, scales = "free") +
  coord_flip() +
  labs(
    title   = "LOO jackknife: fixed-effect estimates",
    subtitle = "Dashed line = full-model estimate",
    x = "Dropped individual", y = "Estimate (log-odds)",
    colour = "Species"
  ) +
  theme_bw(base_size = 9)

print(p_loo)
ggsave("loo_jackknife.png", p_loo, width = 12, height = 8, dpi = 300)

# Summary table: SD of LOO estimates per coefficient (instability measure)
loo_stability <- loo_long |>
  filter(dropped_individual != "FULL MODEL") |>
  group_by(coefficient) |>
  summarise(
    loo_mean = mean(estimate, na.rm = TRUE),
    loo_sd   = sd(estimate, na.rm = TRUE),
    loo_min  = min(estimate, na.rm = TRUE),
    loo_max  = max(estimate, na.rm = TRUE),
    .groups  = "drop"
  ) |>
  left_join(
    pivot_longer(full_coefs[1, ], -c(dropped_individual, dropped_species),
                 names_to = "coefficient", values_to = "full_estimate"),
    by = "coefficient"
  ) 

print(loo_stability)
write.csv(loo_stability, file.path(output_today, "loo_stability_table.csv"), row.names = FALSE)

```



```{r}
# # =============================================================================
# # 4. EFFORT SENSITIVITY CHECK
# # =============================================================================
# # Test whether tracking effort (n_nights, n_positions) correlates with
# # habitat diversity or RSF coefficients, and whether including it as a
# # covariate changes species-level conclusions.
# 
# # --- Correlation: home range size (proxy: n locations) vs habitat diversity ---
# # Replace shannon_H with your actual per-individual diversity measure
# 
# effort_df <- dat |>
#   group_by(individual_id, species) |>
#   summarise(
#     n_used     = sum(used),
#     n_nights   = first(n_nights),
#     # Example diversity: Shannon H on habitat use (adapt to your habitat variable)
#     shannon_H  = {
#       props <- table(cut(forest_cover[used == 1], breaks = 5)) / sum(used)
#       props <- props[props > 0]
#       -sum(props * log(props))
#     },
#     .groups = "drop"
#   )
# 
# p_effort <- ggplot(effort_df,
#                     aes(x = n_nights, y = shannon_H, colour = species)) +
#   geom_point(size = 3) +
#   geom_smooth(method = "lm", se = TRUE) +
#   labs(
#     title = "Tracking effort vs habitat diversity by individual",
#     x     = "Number of nights tracked",
#     y     = "Shannon H (habitat diversity)"
#   ) +
#   theme_bw()
# 
# print(p_effort)
# ggsave("effort_vs_diversity.png", p_effort, width = 6, height = 4, dpi = 300)
# 
# # Formal test per species
# effort_df |>
#   group_by(species) |>
#   summarise(
#     cor_r = cor(n_nights, shannon_H, use = "complete.obs"),
#     cor_p = cor.test(n_nights, shannon_H)$p.value,
#     .groups = "drop"
#   ) |>
#   print()
# 
# # --- Sensitivity model with effort covariate ---
# m_effort <- glmer(
#   used ~ forest_cover + edge_density + open_habitat +
#          species:forest_cover + species:edge_density + species:open_habitat +
#          n_nights +
#          (1 | individual_id),
#   data   = dat,
#   family = binomial(link = "logit"),
#   control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
# )
# 
# # Compare key coefficients between base and effort-adjusted model
# coef_compare <- data.frame(
#   coefficient = names(fixef(m_full)),
#   full_model  = fixef(m_full),
#   effort_model = fixef(m_effort)[names(fixef(m_full))]
# )
# print(coef_compare)
# 
# =============================================================================
# 5. AMONG-INDIVIDUAL VARIANCE BY SPECIES
# =============================================================================


## !! Tried this but encountered errors with the loop that I couldn't fix. 


# Fit separate models per species to compare random-effect (among-individual)
# variance in habitat use.

species_list <- levels(dat$species)

var_by_species <- lapply(species_list, function(sp) {
  m_sp <- glmer(
    used ~  gap3.mean.area + D0 + Hsd + water.distance + (1 | individual_id),
    data   = filter(dat, species == sp),
    family = binomial(link = "logit"),
    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
  )
  vc <- as.data.frame(VarCorr(m_sp))
  data.frame(
    species          = sp,
    re_variance      = vc$vcov[vc$grp == "individual_id"],
    re_sd            = vc$sdcor[vc$grp == "individual_id"]
  )
}) |> bind_rows()


# 
print(var_by_species)
## Random effect variance 
#        species re_variance     re_sd
# 1   M.brandtii  0.01044946 0.1022226
# 2 M.mystacinus  0.02627398 0.1620925

## the random-effect variance appears somewhat larger for M.mystacinus, but both species show relatively low random-effect variability overall.

# Higher RE variance = more among-individual variation in habitat use.
# If M. mystacinus RE variance is LOW, this supports the species-level
# strategy interpretation (all individuals behave similarly) rather than
# individual specialisation.
  
  
  ## Alternative approach
btemp <- dat %>% dplyr::filter(species == "M.brandtii") %>% droplevels() 
  
MBRAfit2 =  glmer(used ~
                   gap3.mean.area +
                   D0 + Hsd + water.distance +
                   (1|BatID), data= btemp,
               family=binomial(link="logit"),nAGQ = 0)

summary(MBRAfit2)

# Generalized linear mixed model fit by maximum likelihood (Adaptive Gauss-Hermite Quadrature, nAGQ = 0) ['glmerMod']
#  Family: binomial  ( logit )
# Formula: used ~ gap3.mean.area + D0 + Hsd + water.distance + (1 | BatID)
#    Data: btemp
# 
#       AIC       BIC    logLik -2*log(L)  df.resid 
#    4344.3    4385.9   -2166.1    4332.3      7540 
# 
# Scaled residuals: 
#     Min      1Q  Median      3Q     Max 
# -0.7425 -0.3564 -0.2698 -0.2009  7.5026 
# 
# Random effects:
#  Groups Name        Variance Std.Dev.
#  BatID  (Intercept) 0.01044  0.1022  
# Number of obs: 7546, groups:  BatID, 9
# 
# Fixed effects:
#                Estimate Std. Error z value Pr(>|z|)    
# (Intercept)    -2.36362    0.05696 -41.495  < 2e-16 ***
# gap3.mean.area -0.62611    0.07141  -8.768  < 2e-16 ***
# D0             -0.45869    0.07141  -6.423 1.33e-10 ***
# Hsd             0.25445    0.04609   5.520 3.39e-08 ***
# water.distance -0.48900    0.04925  -9.930  < 2e-16 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Correlation of Fixed Effects:
#             (Intr) gp3.m. D0     Hsd   
# gap3.mean.r  0.050                     
# D0           0.112  0.733              
# Hsd         -0.158  0.236  0.068       
# water.dstnc  0.180  0.187 -0.029  0.234

mtemp <- dat %>% dplyr::filter(species == "M.mystacinus") %>% droplevels() 
  
MMYSfit2 =  glmer(used ~
                   gap3.mean.area +
                   D0 + Hsd + water.distance +
                   (1|BatID), data= mtemp,
               family=binomial(link="logit"),nAGQ = 0)

summary(MMYSfit2)
# Generalized linear mixed model fit by maximum likelihood (Adaptive Gauss-Hermite Quadrature, nAGQ = 0) ['glmerMod']
#  Family: binomial  ( logit )
# Formula: used ~ gap3.mean.area + D0 + Hsd + water.distance + (1 | BatID)
#    Data: mtemp
# 
#       AIC       BIC    logLik -2*log(L)  df.resid 
#    5222.2    5264.8   -2605.1    5210.2      8970 
# 
# Scaled residuals: 
#     Min      1Q  Median      3Q     Max 
# -0.6429 -0.3573 -0.2834 -0.2012  7.7331 
# 
# Random effects:
#  Groups Name        Variance Std.Dev.
#  BatID  (Intercept) 0.02626  0.1621  
# Number of obs: 8976, groups:  BatID, 12
# 
# Fixed effects:
#                Estimate Std. Error z value Pr(>|z|)    
# (Intercept)    -2.57399    0.06957 -37.000   <2e-16 ***
# gap3.mean.area -0.18833    0.06983  -2.697   0.0070 ** 
# D0              0.17088    0.06580   2.597   0.0094 ** 
# Hsd             0.39581    0.04697   8.428   <2e-16 ***
# water.distance -0.54126    0.05174 -10.462   <2e-16 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Correlation of Fixed Effects:
#             (Intr) gp3.m. D0     Hsd   
# gap3.mean.r  0.160                     
# D0          -0.031  0.730              
# Hsd         -0.159  0.235  0.122       
# water.dstnc  0.273  0.043 -0.253  0.126


vc <- as.data.frame(VarCorr(MBRAfit2))
  data.frame(
    species          = sp,
    re_variance      = vc$vcov[vc$grp == "individual_id"],
    re_sd            = vc$sdcor[vc$grp == "individual_id"]
  ) |> bind_rows()


```



```{r}

```


## Compare home ranges

"A Mann-Whitney U test (sometimes called the Wilcoxon rank-sum test) is used to compare the differences between two independent samples when the sample distributions are not normally distributed and the sample sizes are small (n <30).

It is considered to be the nonparametric equivalent to the two-sample independent t-test." 
# https://www.statology.org/mann-whitney-u-test-r/ 

```{r}
HR <- read_excel("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/HomerangeOverview.xlsx", 
     col_types = c("text", "text", "text", 
         "numeric", "text", "numeric", "numeric", 
         "text", "numeric", "numeric"))
HR$Species <- as.factor(HR$Species)
HR$BatID <- as.factor(HR$BatID)
mbrahr <- HR %>% 
  filter(Species == "M.brandtii") %>% 
  droplevels() %>% 
  select(HRarea) 
mbrahr <- mbrahr$HRarea

mmyshr <- HR %>% 
  filter(Species == "M.mystacinus") %>% 
  droplevels() %>% 
  select(HRarea) 
mmyshr <- mmyshr$HRarea
# 

summary(mbrahr)
 # Min.   : 0.110  
 # 1st Qu.: 1.600  
 # Median : 1.940  
 # Mean   : 3.999  
 # 3rd Qu.: 5.230  
 # Max.   :12.740  
summary(mmyshr)
 # Min.   :0.0800  
 # 1st Qu.:0.3675  
 # Median :0.5150  
 # Mean   :0.7175  
 # 3rd Qu.:1.0975  
 # Max.   :1.8700  
# 	Wilcoxon rank sum exact test

wilcox.test(mbrahr, mmyshr)
# 
# data:  mbrahr and mmyshr
# W = 92, p-value = 0.00562
# alternative hypothesis: true location shift is not equal to 0
# There is strong evidence that the home range size between the two species is distinct from each other. 
```

