---
title: "Quality control and exploration"
output: html_document
date: "2026-06-15"
---

```{r}
### Prepare the work space
library(tidyverse)
library(dplyr)
library(scales)
library(beepr)
library(lubridate)
library(ResourceSelection)
library(RColorBrewer)
library(cowplot)
library(colorBlindness)
library(lme4)
library(stats)
library(DHARMa)
library(lattice)
library(broom)
library(modelsummary)
library(gt)
library(readxl)
library(xlsx)
library(corrplot)
library(vegan)
library(mapview)
library(sf)
library(influence.ME)
library(raster)

getwd()
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/Documents/1. PhD_Main/GitHub_link/Nittedal/SpatialAndMorphology_MBRA_MMYS"

output <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs"

file.name <- "QualityControlExploration"

todays_date <- Sys.Date()
 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
dir.name
 
output_today <- dir.name
output_today

dir.create(output_today)
output_today
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs/QualityControlExploration_2026-06-15"


```



```{r}
## Import spatial data - the points and HR for all bats 

Brandts_points <- readRDS("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/SpatialData/Brandts_points_sf.rds")
# 686 of 4

Brandts_ranges <- readRDS("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/SpatialData/Brandts_ranges_sf.rds")
# 9 of 3

Mystacinus_points <- readRDS("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/SpatialData/Mystacinus_points_sf.rds")
# 816 of 4

Mystacinus_ranges <- readRDS("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/SpatialData/Mystacinus_ranges_sf.rds")
# 12 of 3

## Quick visual inspection to make sure that worked 
# shared extent across all layers
# bb1 <- st_bbox(Brandts_ranges)
# bb2 <- st_bbox(Brandts_points)
# bb3 <- st_bbox(Mystacinus_ranges)
# bb4 <- st_bbox(Mystacinus_points)
# 
# xmin <- min(bb1["xmin"], bb2["xmin"], bb3["xmin"], bb4["xmin"], na.rm = TRUE)
# xmax <- max(bb1["xmax"], bb2["xmax"], bb3["xmax"], bb4["xmax"], na.rm = TRUE)
# ymin <- min(bb1["ymin"], bb2["ymin"], bb3["ymin"], bb4["ymin"], na.rm = TRUE)
# ymax <- max(bb1["ymax"], bb2["ymax"], bb3["ymax"], bb4["ymax"], na.rm = TRUE)
# 
# xpad <- (xmax - xmin) * 0.08
# ypad <- (ymax - ymin) * 0.08
# 
# common_bbox <- st_bbox(c(
#   xmin = xmin - xpad,
#   xmax = xmax + xpad,
#   ymin = ymin - ypad,
#   ymax = ymax + ypad
# ), crs = st_crs(Brandts_ranges))
# 
# 
# CombinedHR <-
#   tm_shape(Mystacinus_ranges, bbox = common_bbox) +
#   tm_borders(col = "#011f4b", lwd = 2, alpha = 0.7) +
#   tm_fill(col = "#011f4b", alpha = 0.1) +
#   
#   tm_shape(Brandts_ranges) +
#   tm_borders(fill_alpha=.7, col = "#32a5ff", lwd = 2) + 
#   tm_fill(fill_alpha=.1, fill = "#32a5ff") +
#   
#   tm_shape(Mystacinus_points) +
#   tm_dots(col = "#011f4b", size = 0.3, alpha = 0.8) +
#   
#   tm_shape(Brandts_points) +
#   tm_dots(col = "#32a5ff", size = 0.3, alpha = 0.8) +
#   
#   tm_basemap("CartoDB.Voyager") +
#   tm_scalebar(position = c("top", "right"), width = 200, text.size = 1.5)
# 
# CombinedHR

## looks good 

```

## Import RSF data, re-run best model and run quality control tests on the robustness of the model 
```{r}

## Covariates from LiDAR data prepared by Hans Ole 
vars <- readRDS("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/SpatialData/231220_vars_dist50_v4_na-rm.rds")

#Used and available positions - just for referencing. 
rsf.data <- read_csv("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs/HomeRangeExploration_2023-09-28/used_available_pts_combined95KE.csv", 
      col_types = cols(...1 = col_skip()))

rsf.data$BatID <- factor(rsf.data$BatID)
used <- rsf.data %>% filter(used == 1) 

summary(used$BatID)

summary(vars)
str(vars) # sf data frame
# 16522 observations of 47 variables 

names(vars)

dt <- vars
dt$BatID <- factor(dt$BatID) # 21 bats 
dt1df<-st_drop_geometry(dt)
dt1df$Species <- factor(dt1df$Species)
dt1df <- as.data.frame(dt)
dt1df <- dt1df %>% select(-geom)
dim(dt1df)
# 16522 obs of 50 vars

levels(dt$BatID )

## Fix error - Turid should be Mystacinus
dtx <- dt1df %>% mutate(
  UID = factor(case_when(
  BatID == "Amelia" ~ "M01",
  BatID == "Aricia" ~ "M11", 
  BatID == "Astrid" ~ "B07", 
  BatID == "Dagny" ~ "B09", 
  BatID == "Daisy" ~ "M06", 
  BatID == "Ethel" ~ "M08", 
  BatID == "Ida" ~ "M10", 
  BatID == "Kaja" ~ "B01", 
  BatID == "Line" ~ "M05", 
  BatID == "Louise" ~ "M12", 
  BatID == "Maren" ~ "B05", 
  BatID == "Marie" ~ "M02",
  BatID == "Nora" ~ "B03", 
  BatID == "Phoebe" ~ "B06", 
  BatID == "Ragnhild" ~ "M09", 
  BatID == "Reeda" ~ "M04", 
  BatID == "Sofia" ~ "B04", 
  BatID == "Steffi" ~ "B08", 
  BatID == "Stine" ~ "M07", 
  BatID == "Thea"  ~ "B02", 
  BatID == "Turid" ~ "M03")),
  species = factor(case_when(
    BatID == "Turid" ~ "M.mystacinus", 
    TRUE ~ Species
  ))) %>% dplyr::select(-Species) %>% 
  mutate(Species = species) %>% 
  dplyr::select(-species)

test <- dtx %>% dplyr::select(BatID, Species, UID) %>% distinct() %>% arrange(UID)

test ## Fixed! 

summary(dtx)

table(dtx$BatID)
#   Amelia   Aricia   Astrid    Dagny    Daisy    Ethel      Ida 
#      176      990      891      913     1474     1287     1309 
#     Kaja     Line   Louise    Maren    Marie     Nora   Phoebe 
#     1001      605      704     1551      341      539      682 
# Ragnhild    Reeda    Sofia   Steffi    Stine     Thea    Turid 
#      308      770      429      429      924     1111       88

 
 ## Fixed! 


## There is an -Inf value for Hmax which can be corrrected like this: 
dtx$Hmax <- gsub("-Inf", "0.5", dtx$Hmax)
dtx$Hmax <- as.numeric(dtx$Hmax)
summary(dtx)

dt2df <- dtx 
dim(dt2df)
# 16522 obs of 50 vars

# Scale all continuous numerical data 
dt1 <- dt2df 
dt1[,6:45]=scale(dt1[,6:45], scale=TRUE) #dt)

dim(dt1)
# 16522 obs of 50 vars 

```

## Run RSF
```{r}
bigfit2 =  glmer(used ~  Species + 
                   gap3.mean.area + 
                   D0 + Hsd + water.distance +
                   Species*gap3.mean.area + 
                   Species*D0 + 
                   Species*Hsd + 
                   (1|BatID), data=dt1,
               family=binomial(link="logit"),nAGQ = 0)

summary(bigfit2)
# Generalized linear mixed model fit by maximum likelihood (Adaptive Gauss-Hermite Quadrature, nAGQ = 0) ['glmerMod']
#  Family: binomial  ( logit )
# Formula: used ~ Species + gap3.mean.area + D0 + Hsd + water.distance +  
#     Species * gap3.mean.area + Species * D0 + Species * Hsd +      (1 | BatID)
#    Data: dt1
# 
#       AIC       BIC    logLik -2*log(L)  df.resid 
#    9563.3    9640.4   -4771.6    9543.3     16512 
# 
# Scaled residuals: 
#     Min      1Q  Median      3Q     Max 
# -0.7635 -0.3560 -0.2783 -0.2008  7.8062 
# 
# Random effects:
#  Groups Name        Variance Std.Dev.
#  BatID  (Intercept) 0.0187   0.1368  
# Number of obs: 16522, groups:  BatID, 21
# 
# Fixed effects:
#                                    Estimate Std. Error z value Pr(>|z|)    
# (Intercept)                        -2.37265    0.06453 -36.768  < 2e-16 ***
# SpeciesM.mystacinus                -0.19439    0.08878  -2.189   0.0286 *  
# gap3.mean.area                     -0.62317    0.07155  -8.709  < 2e-16 ***
# D0                                 -0.45490    0.07149  -6.364 1.97e-10 ***
# Hsd                                 0.25243    0.04578   5.514 3.51e-08 ***
# water.distance                     -0.51682    0.03585 -14.418  < 2e-16 ***
# SpeciesM.mystacinus:gap3.mean.area  0.43776    0.09973   4.390 1.14e-05 ***
# SpeciesM.mystacinus:D0              0.61467    0.09615   6.393 1.63e-10 ***
# SpeciesM.mystacinus:Hsd             0.14392    0.06491   2.217   0.0266 *  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Correlation of Fixed Effects:
#             (Intr) SpcsM. gp3.m. D0     Hsd    wtr.ds SM.:3. SM.:D0
# SpcsM.mystc -0.709                                                 
# gap3.mean.r  0.030 -0.003                                          
# D0           0.100 -0.076  0.737                                   
# Hsd         -0.157  0.138  0.231  0.074                            
# water.dstnc  0.126  0.054  0.127 -0.022  0.167                     
# SpcsM.m:3.. -0.019  0.084 -0.715 -0.529 -0.162 -0.070              
# SpcsM.my:D0 -0.090  0.070 -0.563 -0.741 -0.075 -0.104  0.757       
# SpcsM.mys:H  0.119 -0.204 -0.155 -0.054 -0.695 -0.053  0.227  0.122

library(performance)

r2_nakagawa(bigfit2)
  # Conditional R2: 0.134
  #    Marginal R2: 0.129
```


## Now re-run the RSF with a random slope instead of a random intercept for Bat ID
```{r}

# Try random slope with D0 
bigfit_d0<- lme4::glmer(
  used ~ Species + 
    gap3.mean.area + 
    D0 + Hsd + water.distance +
    Species * gap3.mean.area + 
    Species * D0 + 
    Species * Hsd + 
    (1 + D0 | BatID),
  data = dt1,
  family = binomial(link = "logit"),
  nAGQ = 0
)

# Try random slope with the gap variable 
bigfit_gap<- lme4::glmer(
  used ~ Species + 
    gap3.mean.area + 
    D0 + Hsd + water.distance +
    Species * gap3.mean.area + 
    Species * D0 + 
    Species * Hsd + 
    (1 +  gap3.mean.area | BatID),
  data = dt1,
  family = binomial(link = "logit"),
  nAGQ = 0
)

# Try random slope with the distance to water variable
bigfit_H20<- lme4::glmer(
  used ~ Species + 
    gap3.mean.area + 
    D0 + Hsd + water.distance +
    Species * gap3.mean.area + 
    Species * D0 + 
    Species * Hsd + 
    (1 +  water.distance | BatID),
  data = dt1,
  family = binomial(link = "logit"),
  nAGQ = 0
)


BIC(bigfit2, bigfit_d0, bigfit_gap, bigfit_H20)
#            df      BIC
# bigfit2    10 9640.403
# bigfit_d0  12 9605.394
# bigfit_gap 12 9566.140
# bigfit_H20 12 9297.663

## Lower BIC = better (context still matters) 

bic_tab <- BIC(bigfit2, bigfit_d0, bigfit_gap, bigfit_H20)
bic_tab$delta_BIC <- bic_tab$BIC - min(bic_tab$BIC)
bic_tab[order(bic_tab$BIC), ]

#            df      BIC delta_BIC
# bigfit_H20 12 9297.663    0.0000
# bigfit_gap 12 9566.140  268.4768
# bigfit_d0  12 9605.394  307.7309
# bigfit2    10 9640.403  342.7401

# Common rough guidelines:
# 
# 0–2: little difference in support
# 2–6: some evidence against the larger-BIC model
# 6–10: strong evidence against it
# >10: very strong evidence against it
# So usually you choose the model with the lowest BIC, then interpret how much worse the others are via ΔBIC.

## Compare H20 model with OG big fit

anova(bigfit2, bigfit_H20, test = "Chisq")
#            npar    AIC    BIC  logLik -2*log(L)  Chisq Df Pr(>Chisq)    
# bigfit2      10 9563.3 9640.4 -4771.6    9543.3                         
# bigfit_H20   12 9205.1 9297.7 -4590.6    9181.1 362.17  2  < 2.2e-16 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

## Treating distance to water as a random slope has a significant improvement on the model 
# 
# 
# Compared with bigfit2, the model bigfit_H20 had substantially lower BIC (9297.7 vs. 9640.4) and provided a significantly better fit to the data (likelihood ratio test: χ² = 362.17, df = 2, p < 2.2 × 10⁻¹⁶). Thus, bigfit_H20 was strongly supported over bigfit2.
```



## Testing individual variation's influnece on RSF results (bigfit2) and home range sizes. 
```{r}
# =============================================================================
#  AMONG-INDIVIDUAL VARIANCE BY SPECIES
# =============================================================================
dat <- dt1 %>% mutate(individual_id = BatID, 
                       species = Species) 

# Fit separate models per species to compare random-effect (among-individual)
# variance in habitat use.

species_list <- levels(dat$species)

var_by_species <- lapply(species_list, function(sp) {
  m_sp <- glmer(
    used ~  gap3.mean.area + D0 + Hsd + water.distance + (1 | individual_id),
    data   = filter(dat, species == sp),
    family = binomial(link = "logit"),
    control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
  )
  vc <- as.data.frame(VarCorr(m_sp))
  data.frame(
    species          = sp,
    re_variance      = vc$vcov[vc$grp == "individual_id"],
    re_sd            = vc$sdcor[vc$grp == "individual_id"]
  )
}) |> bind_rows()


# 
print(var_by_species)
## Random effect variance 
#        species re_variance     re_sd
# 1   M.brandtii  0.01044946 0.1022226
# 2 M.mystacinus  0.02627398 0.1620925

## the random-effect variance appears somewhat larger for M.mystacinus, but both species show relatively low random-effect variability overall.

# Higher RE variance = more among-individual variation in habitat use.
# If M. mystacinus RE variance is LOW, this supports the species-level
# strategy interpretation (all individuals behave similarly) rather than
# individual specialisation.
  
  



```


## Further troubleshoot the model to test the robustness
How much does a single individual influence the outcome of the model? 
```{r}
# =============================================================================
#  COOK'S DISTANCE AND DFBETAS (influence.ME)
# =============================================================================
# This refits the model n times, dropping one individual per iteration.
# Can be slow for large datasets; set parallel = TRUE if using a cluster.

infl <- influence(bigfit2, group = "BatID")
cd <- cooks.distance(infl)

cd_df <- data.frame(
  BatID = rownames(cd),
  cooks_distance = as.numeric(cd)
)

cd_df

#       BatID cooks_distance
# 1    Amelia    0.087254669
# 2    Aricia    0.466942487
# 3    Astrid    0.278351495
# 4     Dagny    0.645118904
# 5     Daisy    0.302865800
# 6     Ethel    0.644116758
# 7       Ida    0.329554532
# 8      Kaja    0.116871787
# 9      Line    0.210178012
# 10   Louise    0.518241172
# 11    Maren    1.476387292
# 12    Marie    0.173244284
# 13     Nora    0.156122941
# 14   Phoebe    0.179989711
# 15 Ragnhild    0.121511521
# 16    Reeda    0.311126351
# 17    Sofia    0.213415269
# 18   Steffi    0.151305957
# 19    Stine    0.532828146
# 20     Thea    4.550412472
# 21    Turid    0.007731942

               
        summary(dt1)

# --- Cook's distance ---
cook_d <- cooks.distance(infl)
cook_df <- data.frame(
  individual_id = rownames(cook_d),
  cooks_d       = as.numeric(cook_d),
  species       = dt1 |>
    distinct(BatID, Species) |>
    arrange(BatID) |>
    pull(Species)
)

cook_df


# Threshold: common rule of thumb is 4/n (conservative) or visual inspection
threshold_cook <- 4 / 21

p_cook <- ggplot(cook_df, aes(x = reorder(individual_id, cooks_d),
                               y = cooks_d, fill = species)) +
  geom_col() +
  geom_hline(yintercept = threshold_cook, linetype = "dashed", colour = "red") +
  coord_flip() +
  labs(
    title = "Cook's distance by individual",
    subtitle = paste0("Dashed line = 4/n threshold (", round(threshold_cook, 3), ")"),
    x = NULL, y = "Cook's D"
  ) +
  theme_bw()


print(p_cook)
#ggsave("cooks_distance.png", p_cook, width = 7, height = 5, dpi = 300)


temp <- dt1 %>% mutate(individual_id = BatID, 
                       species = Species) 
# --- DFBETAS ---
dfb <- dfbetas(infl)
dfb_df <- as.data.frame(dfb) |>
  mutate(individual_id = rownames(dfb)) |>
  pivot_longer(-individual_id, names_to = "coefficient", values_to = "dfbeta") |>
  left_join(distinct(temp, individual_id, species), by = "individual_id")

# DFBETAS threshold: |DFBETAS| > 2/sqrt(n)
threshold_dfb <- 2 / sqrt(21)

p_dfb <- ggplot(dfb_df, aes(x = reorder(individual_id, abs(dfbeta)),
                              y = dfbeta, colour = species)) +
  geom_point(size = 2) +
  geom_hline(yintercept =  threshold_dfb, linetype = "dashed", colour = "red") +
  geom_hline(yintercept = -threshold_dfb, linetype = "dashed", colour = "red") +
  facet_wrap(~ coefficient, scales = "free_y") +
  coord_flip() +
  labs(
    title = "DFBETAS by individual and coefficient",
    subtitle = paste0("Dashed lines = ±2/√n threshold (±", round(threshold_dfb, 3), ")"),
    x = NULL, y = "DFBETAS"
  ) +
  theme_bw(base_size = 9)

print(p_dfb)
#ggsave("dfbetas.png", p_dfb, width = 10, height = 7, dpi = 300)


# =============================================================================
#  LEAVE-ONE-OUT (LOO) JACKKNIFE BY INDIVIDUAL
# =============================================================================
# Refit model dropping one individual at a time; record fixed-effect estimates.

# bigfit2 =  glmer(used ~  Species + 
#                    gap3.mean.area + 
#                    D0 + Hsd + water.distance +
#                    Species*gap3.mean.area + 
#                    Species*D0 + 
#                    Species*Hsd + 
#                    (1|BatID), data=dt1,
#                family=binomial(link="logit"),nAGQ = 0)


str(temp)
dat <- temp

individuals <- unique(dat$individual_id)

loo_results <- lapply(individuals, function(id) {
  dat_sub <- filter(dat, individual_id != id)
  m_loo <- tryCatch(
    glmer(
      used ~ Species + gap3.mean.area + 
                    D0 + Hsd + water.distance +
                    Species*gap3.mean.area + 
                    Species*D0 + 
                    Species*Hsd + 
             (1 | individual_id),
      data   = dat_sub,
      family = binomial(link = "logit"),
      control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
    ),
    error = function(e) NULL
  )
  if (is.null(m_loo)) return(NULL)
  coef_df <- as.data.frame(t(fixef(m_loo)))
  coef_df$dropped_individual <- id
  coef_df$dropped_species    <- dat |>
    filter(individual_id == id) |>
    pull(species) |>
    unique()
  coef_df
}) |>
  bind_rows()

# Full-model estimates for reference line
full_coefs <- data.frame(t(fixef(bigfit2)))
full_coefs$dropped_individual <- "FULL MODEL"
full_coefs$dropped_species    <- NA

# Combine and reshape
loo_long <- bind_rows(loo_results, full_coefs) |>
  pivot_longer(
    cols      = -c(dropped_individual, dropped_species),
    names_to  = "coefficient",
    values_to = "estimate"
  )

# Plot: one panel per coefficient, each point is the LOO estimate
p_loo <- loo_long |>
  filter(dropped_individual != "FULL MODEL") |>
  ggplot(aes(x = reorder(dropped_individual, estimate),
             y = estimate, colour = dropped_species)) +
  geom_point(size = 2) +
  geom_hline(
    data = filter(loo_long, dropped_individual == "FULL MODEL"),
    aes(yintercept = estimate),
    linetype = "dashed", colour = "black"
  ) +
  facet_wrap(~ coefficient, scales = "free") +
  coord_flip() +
  labs(
    title   = "LOO jackknife: fixed-effect estimates",
    subtitle = "Dashed line = full-model estimate",
    x = "Dropped individual", y = "Estimate (log-odds)",
    colour = "Species"
  ) +
  theme_bw(base_size = 9)

print(p_loo)
# 1: Removed 84 rows containing missing values or values outside the scale range (`geom_point()`). 
# 2: Removed 4 rows containing missing values or values outside the scale range (`geom_hline()`). 
#ggsave("loo_jackknife.png", p_loo, width = 12, height = 8, dpi = 300)

# Summary table: SD of LOO estimates per coefficient (instability measure)
loo_stability <- loo_long |>
  filter(dropped_individual != "FULL MODEL") |>
  group_by(coefficient) |>
  summarise(
    loo_mean = mean(estimate, na.rm = TRUE),
    loo_sd   = sd(estimate, na.rm = TRUE),
    loo_min  = min(estimate, na.rm = TRUE),
    loo_max  = max(estimate, na.rm = TRUE),
    .groups  = "drop"
  ) |>
  left_join(
    pivot_longer(full_coefs[1, ], -c(dropped_individual, dropped_species),
                 names_to = "coefficient", values_to = "full_estimate"),
    by = "coefficient"
  ) 

print(loo_stability)
#write.csv(loo_stability, file.path(output_today, "loo_stability_table.csv"), row.names = FALSE)

################################################################################
######################### Separate for each species ############################
################################################################################
# A bit difficult to interpret when the two species are together - I will try again with separate models for each species

mbra <- dt1 %>% dplyr::filter(Species == "M.brandtii") %>% droplevels()
mmys <- dt1 %>% dplyr::filter(Species == "M.mystacinus") %>% droplevels()

##############################################
## MBRA ## 

mbrafit =  glmer(used ~  
                   gap3.mean.area +
                   D0 + Hsd + water.distance +
                   (1|BatID), data=mbra,
               family=binomial(link="logit"),nAGQ = 0)

summary(mbrafit)
#       AIC       BIC    logLik -2*log(L)  df.resid 
#    4344.3    4385.9   -2166.1    4332.3      7540 
# 
# Scaled residuals: 
#     Min      1Q  Median      3Q     Max 
# -0.7425 -0.3564 -0.2698 -0.2009  7.5026 
# 
# Random effects:
#  Groups Name        Variance Std.Dev.
#  BatID  (Intercept) 0.01044  0.1022  
# Number of obs: 7546, groups:  BatID, 9
# 
# Fixed effects:
#                Estimate Std. Error z value Pr(>|z|)    
# (Intercept)    -2.36362    0.05696 -41.495  < 2e-16 ***
# gap3.mean.area -0.62611    0.07141  -8.768  < 2e-16 ***
# D0             -0.45869    0.07141  -6.423 1.33e-10 ***
# Hsd             0.25445    0.04609   5.520 3.39e-08 ***
# water.distance -0.48900    0.04925  -9.930  < 2e-16 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Correlation of Fixed Effects:
#             (Intr) gp3.m. D0     Hsd   
# gap3.mean.r  0.050                     
# D0           0.112  0.733              
# Hsd         -0.158  0.236  0.068       
# water.dstnc  0.180  0.187 -0.029  0.234


# LOO
bats <- unique(mbra$BatID)

# Leave-one-individual-out models
loo_results <- lapply(bats, function(id) {
  
  mbra_sub <- mbra %>%
    filter(BatID != id)
  
  m_loo <- tryCatch(
    glmer(
      used ~ gap3.mean.area + D0 + Hsd + water.distance + (1 | BatID),
      data = mbra_sub,
      family = binomial(link = "logit"),
      nAGQ = 0
    ),
    error = function(e) NULL
  )
  
  if (is.null(m_loo)) return(NULL)
  
  coef_df <- as.data.frame(t(fixef(m_loo)))
  coef_df$dropped_BatID <- id
  coef_df
}) %>%
  bind_rows()


# Add the full-model coefficients for comparison
full_coefs <- as.data.frame(t(fixef(mbrafit)))
full_coefs$dropped_BatID <- "FULL MODEL"

# Reshape to long format
loo_long <- bind_rows(loo_results, full_coefs) %>%
  pivot_longer(
    cols = -dropped_BatID,
    names_to = "coefficient",
    values_to = "estimate"
  )

# Plot the influence of dropping each individual
p_loo <- loo_long %>%
  filter(dropped_BatID != "FULL MODEL") %>%
  ggplot(aes(x = reorder(dropped_BatID, estimate), y = estimate)) +
  geom_point(size = 2) +
  geom_hline(
    data = loo_long %>% filter(dropped_BatID == "FULL MODEL"),
    aes(yintercept = estimate),
    linetype = "dashed",
    colour = "black"
  ) +
  facet_wrap(~ coefficient, scales = "free") +
  coord_flip() +
  labs(
    title = "Leave-one-individual-out jackknife",
    subtitle = "Dashed line = full-model estimate",
    x = "Dropped BatID",
    y = "Estimate (log-odds)"
  ) +
  theme_bw()

print(p_loo)


# Summary table of coefficient stability

loo_stability <- loo_long %>%
  filter(dropped_BatID != "FULL MODEL") %>%
  group_by(coefficient) %>%
  summarise(
    loo_mean = mean(estimate, na.rm = TRUE),
    loo_sd   = sd(estimate, na.rm = TRUE),
    loo_min  = min(estimate, na.rm = TRUE),
    loo_max  = max(estimate, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  left_join(
    full_coefs %>%
      pivot_longer(
        cols = -dropped_BatID,
        names_to = "coefficient",
        values_to = "full_estimate"
      ),
    by = "coefficient"
  )

# print(loo_stability)
#   coefficient    loo_mean loo_sd loo_min loo_max dropped_BatID full_estimate
#   <chr>             <dbl>  <dbl>   <dbl>   <dbl> <chr>                 <dbl>
# 1 (Intercept)      -2.38  0.107   -2.66   -2.31  FULL MODEL           -2.36 
# 2 D0               -0.460 0.146   -0.720  -0.160 FULL MODEL           -0.459
# 3 Hsd               0.253 0.0492   0.137   0.302 FULL MODEL            0.254
# 4 gap3.mean.area   -0.628 0.135   -0.858  -0.411 FULL MODEL           -0.626
# 5 water.distance   -0.506 0.196   -1.01   -0.369 FULL MODEL           -0.489


# SIKT interpretation:

# The leave-one-individual-out analysis indicates that fixed-effect estimates are generally robust to the removal of any single bat. The direction of all effects remained unchanged across LOO runs, suggesting qualitative inference is stable. Among predictors, water.distance showed the greatest sensitivity to individual removal, whereas Hsd and the intercept were most stable.


### MMYS ###

##############################################
## MBRA ## 

mmysfit =  glmer(used ~  
                   gap3.mean.area +
                   D0 + Hsd + water.distance +
                   (1|BatID), data=mmys,
               family=binomial(link="logit"),nAGQ = 0)

summary(mmysfit)
#       AIC       BIC    logLik -2*log(L)  df.resid 
#    5222.2    5264.8   -2605.1    5210.2      8970 
# 
# Scaled residuals: 
#     Min      1Q  Median      3Q     Max 
# -0.6429 -0.3573 -0.2834 -0.2012  7.7331 
# 
# Random effects:
#  Groups Name        Variance Std.Dev.
#  BatID  (Intercept) 0.02626  0.1621  
# Number of obs: 8976, groups:  BatID, 12
# 
# Fixed effects:
#                Estimate Std. Error z value Pr(>|z|)    
# (Intercept)    -2.57399    0.06957 -37.000   <2e-16 ***
# gap3.mean.area -0.18833    0.06983  -2.697   0.0070 ** 
# D0              0.17088    0.06580   2.597   0.0094 ** 
# Hsd             0.39581    0.04697   8.428   <2e-16 ***
# water.distance -0.54126    0.05174 -10.462   <2e-16 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Correlation of Fixed Effects:
#             (Intr) gp3.m. D0     Hsd   
# gap3.mean.r  0.160                     
# D0          -0.031  0.730              
# Hsd         -0.159  0.235  0.122       
# water.dstnc  0.273  0.043 -0.253  0.126


# LOO
bats <- unique(mmys$BatID)

# Leave-one-individual-out models
loo_results <- lapply(bats, function(id) {
  
  mmys_sub <- mmys %>%
    filter(BatID != id)
  
  m_loo <- tryCatch(
    glmer(
      used ~ gap3.mean.area + D0 + Hsd + water.distance + (1 | BatID),
      data = mmys_sub,
      family = binomial(link = "logit"),
      nAGQ = 0
    ),
    error = function(e) NULL
  )
  
  if (is.null(m_loo)) return(NULL)
  
  coef_df <- as.data.frame(t(fixef(m_loo)))
  coef_df$dropped_BatID <- id
  coef_df
}) %>%
  bind_rows()


# Add the full-model coefficients for comparison
full_coefs <- as.data.frame(t(fixef(mmysfit)))
full_coefs$dropped_BatID <- "FULL MODEL"

# Reshape to long format
loo_long <- bind_rows(loo_results, full_coefs) %>%
  pivot_longer(
    cols = -dropped_BatID,
    names_to = "coefficient",
    values_to = "estimate"
  )

# Plot the influence of dropping each individual
p_loo <- loo_long %>%
  filter(dropped_BatID != "FULL MODEL") %>%
  ggplot(aes(x = reorder(dropped_BatID, estimate), y = estimate)) +
  geom_point(size = 2) +
  geom_hline(
    data = loo_long %>% filter(dropped_BatID == "FULL MODEL"),
    aes(yintercept = estimate),
    linetype = "dashed",
    colour = "black"
  ) +
  facet_wrap(~ coefficient, scales = "free") +
  coord_flip() +
  labs(
    title = "Leave-one-individual-out jackknife",
    subtitle = "Dashed line = full-model estimate",
    x = "Dropped BatID",
    y = "Estimate (log-odds)"
  ) +
  theme_bw()

print(p_loo)


# Summary table of coefficient stability

loo_stability <- loo_long %>%
  filter(dropped_BatID != "FULL MODEL") %>%
  group_by(coefficient) %>%
  summarise(
    loo_mean = mean(estimate, na.rm = TRUE),
    loo_sd   = sd(estimate, na.rm = TRUE),
    loo_min  = min(estimate, na.rm = TRUE),
    loo_max  = max(estimate, na.rm = TRUE),
    .groups  = "drop"
  ) %>%
  left_join(
    full_coefs %>%
      pivot_longer(
        cols = -dropped_BatID,
        names_to = "coefficient",
        values_to = "full_estimate"
      ),
    by = "coefficient"
  )

 print(loo_stability)
#   coefficient    loo_mean loo_sd loo_min loo_max dropped_BatID full_estimate
#   <chr>             <dbl>  <dbl>   <dbl>   <dbl> <chr>                 <dbl>
# 1 (Intercept)      -2.58  0.0410 -2.64   -2.51   FULL MODEL           -2.57 
# 2 D0                0.172 0.0636  0.0637  0.276  FULL MODEL            0.171
# 3 Hsd               0.395 0.0398  0.331   0.458  FULL MODEL            0.396
# 4 gap3.mean.area   -0.189 0.0746 -0.328  -0.0970 FULL MODEL           -0.188
# 5 water.distance   -0.544 0.0565 -0.675  -0.478  FULL MODEL           -0.541


# SIKT interpretation:

# The leave-one-individual-out analysis indicates that the model is highly robust to the removal of any single bat. For all predictors, the LOO means were nearly identical to the full-model estimates, and the sign of each coefficient was unchanged across all iterations. Among the covariates, gap3.mean.area and D0 showed somewhat greater variation than the other terms, but overall the model appears very stable.

 
 
```


```{r}

############# Sensitivity checks ####################

### Home range size first 

Brandts_ranges$species <- "M.brandtii"
Mystacinus_ranges$species <- "M.mystacinus"

BHR <- rbind(Brandts_ranges, Mystacinus_ranges)
#21 of 3 - sweet 


temp <- read_excel("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/BatsMeta.xlsx")

names(temp)

effort <- temp %>% 
  dplyr::select(BatID, HRha, Fixes, DaysTracked, Species) %>% 
  rename(n_nights = DaysTracked,
         n_positions = Fixes, 
         area = HRha,
         species = Species) %>% 
  mutate(BatID = factor(BatID),
         species = factor(species)) %>% 
  dplyr::filter(BatID != "Cecilia") %>% 
  droplevels()

summary(effort)

temp <- effort %>% dplyr::select(BatID, area, n_nights, n_positions)

### Now include the HR size and n points as columns to the larger RSF dataset 
dt2 <- left_join(dt1, temp)
summary(dt2)
# Looking good!! 

## Now the effort dataset - with the number of nights and the number of points for each individuals 

stest <- lm(log(area) ~ species + log(n_positions) + log(n_nights), data = effort)
summary(stest)

# Residuals:
#      Min       1Q   Median       3Q      Max 
# -2.77717 -0.57034 -0.06847  0.62102  1.85620 
# 
# Coefficients:
#                     Estimate Std. Error t value Pr(>|t|)  
# (Intercept)           4.0255     1.6737   2.405   0.0278 *
# speciesM.mystacinus  -1.1020     0.5712  -1.929   0.0706 .
# log(n_positions)     -0.3020     0.5020  -0.602   0.5554  
# log(n_nights)         1.2907     0.9302   1.388   0.1832  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 1.165 on 17 degrees of freedom
# Multiple R-squared:  0.3766,	Adjusted R-squared:  0.2665 
# F-statistic: 3.423 on 3 and 17 DF,  p-value: 0.04105

## Alternative 

stest1 <- lm(log(area) ~ species * log(n_positions) + species * log(n_nights), data = effort)

summary(stest1)
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -2.45104 -0.39529 -0.07701  0.46841  1.82507 
# 
# Coefficients:
#                                      Estimate Std. Error t value Pr(>|t|)  
# (Intercept)                            8.8811     4.8998   1.813    0.090 .
# speciesM.mystacinus                   -6.7849     5.1603  -1.315    0.208  
# log(n_positions)                      -1.5001     0.9885  -1.517    0.150  
# log(n_nights)                          1.4053     2.1171   0.664    0.517  
# speciesM.mystacinus:log(n_positions)   1.5835     1.1507   1.376    0.189  
# speciesM.mystacinus:log(n_nights)     -0.5245     2.3707  -0.221    0.828  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 1.163 on 15 degrees of freedom
# Multiple R-squared:  0.4521,	Adjusted R-squared:  0.2694 
# F-statistic: 2.475 on 5 and 15 DF,  p-value: 0.07961

## Higher R value for tht second model and I think that the structure of the model is more ecologically meaningful - with interactions between species and each effor metric. 

## Interpreation - basically, seems to be very little influence overall. 

## Check for collinearity 

cor(effort$n_positions, effort$n_nights)
# 0.6568861 
# That is pretty high ... so I will fit separate models for each effort metric 

stest2 <- lm(log(area) ~ species * log(n_positions), data = effort)

summary(stest2)
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -2.65112 -0.54310  0.09984  0.58401  1.61207 
# 
# Coefficients:
#                                      Estimate Std. Error t value Pr(>|t|)  
# (Intercept)                           10.7955     3.8575   2.799   0.0123 *
# speciesM.mystacinus                   -8.6159     4.1660  -2.068   0.0542 .
# log(n_positions)                      -1.2739     0.9037  -1.410   0.1767  
# speciesM.mystacinus:log(n_positions)   1.7143     0.9841   1.742   0.0996 .
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 1.132 on 17 degrees of freedom
# Multiple R-squared:  0.4111,	Adjusted R-squared:  0.3071 
# F-statistic: 3.955 on 3 and 17 DF,  p-value: 0.02617

stest3 <- lm(log(area) ~ species * log(n_nights), data = effort)
summary(stest3)
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -2.95586 -0.43259 -0.06082  0.61036  1.75639 
# 
# Coefficients:
#                                   Estimate Std. Error t value Pr(>|t|)
# (Intercept)                         4.7741     4.1232   1.158    0.263
# speciesM.mystacinus                -2.5364     4.3189  -0.587    0.565
# log(n_nights)                       0.2979     2.0061   0.148    0.884
# speciesM.mystacinus:log(n_nights)   0.6938     2.1353   0.325    0.749
# 
# Residual standard error: 1.174 on 17 degrees of freedom
# Multiple R-squared:  0.3672,	Adjusted R-squared:  0.2556 
# F-statistic: 3.289 on 3 and 17 DF,  p-value: 0.04613

## Still a low influence for both metrics! '

### Now visualize

ggplot(effort, aes(n_positions, area, color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  xlab("Number of positions") + 
  ylab("Homerange area (hectares)")+ 
  ggtitle("Influence of the number of positions on size of home range") +
  theme_bw()

ggplot(effort, aes(n_nights, area, color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  xlab("Number of nights tracked") + 
  ylab("Homerange area (hectares)")+ 
  ggtitle("Influence of the number of nights tracked on size of home range") +
  theme_bw()

### Interpretation

### More influence on MBRA than on MMYS for both metrics however - this a very wide spread 


### Suggested phrasing from Sikt: 
# We evaluated whether home-range area estimates were sensitive to tracking effort by modeling estimated area as a function of number of locations and number of nights tracked, including species as a covariate. 

# =============================================================================
#   EFFORT SENSITIVITY CHECK
# =============================================================================
# Test whether tracking effort (n_nights, n_positions) correlates with
# habitat diversity or RSF coefficients, and whether including it as a
# covariate changes species-level conclusions.

# --- Correlation: home range size (proxy: n locations) vs habitat diversity ---
# Replace shannon_H with your actual per-individual diversity measure

## Run for each of the main habitat types 
dat <- dt2 %>% mutate(individual_id = BatID,
                      species = Species)

########################## D0 #########################################
effort_D0 <- dat |>
  group_by(individual_id, species) |>
  summarise(
    n_used     = sum(used),
    n_nights   = first(n_nights),
    # Example diversity: Shannon H on habitat use (adapt to your habitat variable)
    shannon_H  = {
      props <- table(cut(D0[used == 1], breaks = 5)) / sum(used)
      props <- props[props > 0]
      -sum(props * log(props))
    },
    .groups = "drop"
  )

p_effort <- ggplot(effort_D0,
                    aes(x = n_nights, y = shannon_H, colour = species)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Tracking effort vs D0 habitat diversity by individual",
    x     = "Number of nights tracked",
    y     = "Shannon H (habitat diversity)"
  ) +
  theme_bw()

print(p_effort)
ggsave(file = file.path(output_today,"effort_vs_diversity_D0.png"), p_effort, width = 6, height = 4, dpi = 300)

# Formal test per species
effort_D0 |>
  group_by(species) |>
  summarise(
    cor_r = cor(n_nights, shannon_H, use = "complete.obs"),
    cor_p = cor.test(n_nights, shannon_H)$p.value,
    .groups = "drop"
  ) |>
  print()

########################## HSD #########################################
effort_Hsd <- dat |>
  group_by(individual_id, species) |>
  summarise(
    n_used     = sum(used),
    n_nights   = first(n_nights),
    # Example diversity: Shannon H on habitat use (adapt to your habitat variable)
    shannon_H  = {
      props <- table(cut(Hsd[used == 1], breaks = 5)) / sum(used)
      props <- props[props > 0]
      -sum(props * log(props))
    },
    .groups = "drop"
  )

p_effort <- ggplot(effort_Hsd,
                    aes(x = n_nights, y = shannon_H, colour = species)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Tracking effort vs Hsd habitat diversity by individual",
    x     = "Number of nights tracked",
    y     = "Shannon H (habitat diversity)"
  ) +
  theme_bw()

print(p_effort)
ggsave(file = file.path(output_today,"effort_vs_diversity_Hsd.png"), p_effort, width = 6, height = 4, dpi = 300)

# Formal test per species
effort_Hsd |>
  group_by(species) |>
  summarise(
    cor_r = cor(n_nights, shannon_H, use = "complete.obs"),
    cor_p = cor.test(n_nights, shannon_H)$p.value,
    .groups = "drop"
  ) |>
  print()

########################## gap3.mean.area #########################################
effort_gap <- dat |>
  group_by(individual_id, species) |>
  summarise(
    n_used     = sum(used),
    n_nights   = first(n_nights),
    # Example diversity: Shannon H on habitat use (adapt to your habitat variable)
    shannon_H  = {
      props <- table(cut(gap3.mean.area[used == 1], breaks = 5)) / sum(used)
      props <- props[props > 0]
      -sum(props * log(props))
    },
    .groups = "drop"
  )

p_effort <- ggplot(effort_gap,
                    aes(x = n_nights, y = shannon_H, colour = species)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Tracking effort vs gap3.mean.area habitat diversity by individual",
    x     = "Number of nights tracked",
    y     = "Shannon H (habitat diversity)"
  ) +
  theme_bw()

print(p_effort)
ggsave(file = file.path(output_today,"effort_vs_diversity_gap3.mean.area.png"), p_effort, width = 6, height = 4, dpi = 300)

effort_gap |>
  group_by(species) |>
  summarise(
    cor_r = cor(n_nights, shannon_H, use = "complete.obs"),
    cor_p = cor.test(n_nights, shannon_H)$p.value,
    .groups = "drop"
  ) |>
  print()


########################## water.distance #########################################
effort_water <- dat |>
  group_by(individual_id, species) |>
  summarise(
    n_used     = sum(used),
    n_nights   = first(n_nights),
    # Example diversity: Shannon H on habitat use (adapt to your habitat variable)
    shannon_H  = {
      props <- table(cut(water.distance[used == 1], breaks = 5)) / sum(used)
      props <- props[props > 0]
      -sum(props * log(props))
    },
    .groups = "drop"
  )

p_effort <- ggplot(effort_water,
                    aes(x = n_nights, y = shannon_H, colour = species)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Tracking effort vs water.distance habitat diversity by individual",
    x     = "Number of nights tracked",
    y     = "Shannon H (habitat diversity)"
  ) +
  theme_bw()

print(p_effort)
ggsave(file = file.path(output_today,"effort_vs_diversity_water.distance.png"), p_effort, width = 6, height = 4, dpi = 300)

effort_water |>
  group_by(species) |>
  summarise(
    cor_r = cor(n_nights, shannon_H, use = "complete.obs"),
    cor_p = cor.test(n_nights, shannon_H)$p.value,
    .groups = "drop"
  ) |>
  print()

################################################################################
# Formal test per species
effort_gap |>
  group_by(species) |>
  summarise(
    cor_r = cor(n_nights, shannon_H, use = "complete.obs"),
    cor_p = cor.test(n_nights, shannon_H)$p.value,
    .groups = "drop"
  ) |>
  print()

# --- Sensitivity model with effort covariate ---
m_effort <- glmer(
 used ~  Species +  gap3.mean.area + 
   D0 + Hsd + water.distance +
   Species*gap3.mean.area + 
   Species*Hsd + 
   n_nights + 
  (1|BatID), data=dt2,
 family=binomial(link="logit"),nAGQ = 0,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)

# Compare key coefficients between base and effort-adjusted model
coef_compare <- data.frame(
  coefficient = names(fixef(bigfit2)),
  full_model  = fixef(bigfit2),
  effort_model = fixef(m_effort)[names(fixef(bigfit2))]
)
print(coef_compare)
# (Intercept)                                               (Intercept) -2.3726510  -2.36257872
# SpeciesM.mystacinus                               SpeciesM.mystacinus -0.1943861  -0.22608964
# gap3.mean.area                                         gap3.mean.area -0.6231665  -0.36312096 *** .3 difference
# D0                                                                 D0 -0.4548980  -0.12822280 *** .3 difference
# Hsd                                                               Hsd  0.2524302   0.27451972
# water.distance                                         water.distance -0.5168179  -0.48966117
# SpeciesM.mystacinus:gap3.mean.area SpeciesM.mystacinus:gap3.mean.area  0.4377611  -0.04761241 *** .4 difference
# SpeciesM.mystacinus:D0                         SpeciesM.mystacinus:D0  0.6146687           NA *** NA?
# SpeciesM.mystacinus:Hsd                       SpeciesM.mystacinus:Hsd  0.1439242   0.09295824
```

## Katrine's further recommendations
GLM
Prob.use ~ D0 + homerange size + D0*homerange size + Species + (1 + BatID), data=dt1,           
family=binomial(link="logit"),nAGQ = 0)

Prob.use ~ Hsd + homerange size + Hsd*homerange size
Prob.use ~ water.distance + homerange size + water.distance*homerange size 
Prob.use ~ gap3.mean + homerange size + gap3.mean*homerange size


### 
RSF version 2 h20
- include HR in the model as a fixed effect
- include HR as an offset variable 
