---
title: "Revisions"
output: html_document
date: "2026-03-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Following up on first round of revisions to per editor's feedback
## March 2025



# Set up the work space
```{r}
##############################################################
library(rlang)
library(cowplot)
library(tidyverse) #
library(readxl) ## 
library(reshape2) 
library(ggpubr) # 
library(cowplot) # 
library(devtools) #
library(phyloseq)
library(plyr)
library(Rcpp)
library(vctrs)
library(sjPlot)
library(effects)
library(ade4)
library(pals)
library(RColorBrewer)
library(MetBrewer)
library(knitr)
library(kableExtra)
library(colorBlindness)
library(devtools)

library(SpatialKDE)
library(tmap)
library(tmaptools)
library(tidyverse)
library(sp)
library(sf)
library(rgeos)
library(adehabitatMA)
library(digest)
library(mime)
library(jsonlite)
library(crs)
library(dplyr)
library(adehabitatHR)
library(Rcpp)
library(raster)
library(rgdal)
library(munsell)
library(tibble)
library(pillar)
library(ggplot2)
library(ResourceSelection)
library(mapview)
library(amt)
library(sjPlot)
library(Rgb)
library(nloptr)
library(colorspace)
library(rlang)
library(lattice)
library(dismo)
library(scales)
library(rJava)
library(beepr)
library(OpenStreetMap)
library(lubridate)
library(rgdal)
library(munsell)
library(pillar)
library(mapview)
library(amt)
library(sjPlot)
library(Rgb)
library(nloptr)
library(rlang)
library(broom) 
library(MASS)
library(ggmap)
library(gridExtra)
library(raster)
library(leaflet)
library(terra)

##############################################################
getwd()
# "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/Documents/1. PhD_Main/GitHub_link/Nittedal/Diet_Myotis_brandts_mystacinus"

output <- "C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs"

file.name <- "Revisions"

todays_date <- Sys.Date()
 
dir.name <- str_c(output,"/", file.name, "_", todays_date)
dir.name
 
output_today <- dir.name
output_today

dir.create(output_today)
output_today


## Input

df <- read_delim("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/batdata.csv", delim =";")

cap <- read_delim("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/capture (2).csv", delim =";")

meta <- read_excel("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/BatsMeta.xlsx")

#157 obs 5 vars 


```



```{r}

names(cap)
tag <- cap %>% 
  dplyr::filter(Name != "") %>% 
  dplyr::select(Date, Species, Sex, `Reprod.`, Age, Weight, 
                `BM1 (g)`, `Forearm l. (mm. )`, 
                `Photo ID`, Name) %>%
  dplyr::mutate(Species = factor(Species), 
                Sex = factor(Sex),
                Name = factor(Name),
                ID = factor(`Photo ID`), 
                RFA = `Forearm l. (mm. )`,
                BM1 = `BM1 (g)`,
                REPRO = `Reprod.`) %>% 
  select(-c(`BM1 (g)`, `Forearm l. (mm. )`, `Photo ID`, `Reprod.` )) %>% 
  arrange(Species, ID)

summary(tag)

temp <- tag %>% select(ID, Name) %>% droplevels()

temp


## Report the average capture weight and RFA of MBRA and MMYS from all captures 
cap1 <- cap %>% 
  dplyr::select(Date, Species, Sex, `Reprod.`, Age, Weight, 
                `BM1 (g)`, `Forearm l. (mm. )`, 
                `Photo ID`, Name) %>%
  dplyr::mutate(Species = factor(Species), 
                Sex = factor(Sex),
                Name = factor(Name),
                ID = factor(`Photo ID`), 
                RFA = `Forearm l. (mm. )`,
                BM1 = `BM1 (g)`,
                REPRO = `Reprod.`,
                Date = as.Date(Date, format = "%d.%m.%Y")) %>% 
  select(-c(`BM1 (g)`, `Forearm l. (mm. )`, `Photo ID`, `Reprod.` )) %>% 
  dplyr::filter(!is.na(Date)) %>% # drop the empty rows 
  arrange(Species, ID) 

# 223 obs of 10 vars 
summary(cap1)
# date error that needs fixing

cap1$Date <- update(cap1$Date, year = 2018)
summary(cap1)

# Drop escaped bats or those that were not fully processed (escapees or released early) 
# 4 cases - 1 MDAU, 1 Myotis spp., 1 PAUR, 1 unknown species 

cap2 <- cap1 %>% dplyr::filter(!is.na(RFA))
#219 of 10 vars

summary(cap2)


levels(cap2$Species)
summary(cap2$Species)

mbra <- cap2 %>% dplyr::filter(Species == "Myotis brandtii") %>% droplevels() 

summary(mbra)
sd(mbra$Weight)

mmys <- cap2 %>% dplyr::filter(Species == "Myotis mystacinus") %>% droplevels() 

summary(mmys)
sd(mmys$Weight)
```


## Convert HR from M2 to HA and re-run the Mann Whitney U test 
## Compare home ranges

"A Mann-Whitney U test (sometimes called the Wilcoxon rank-sum test) is used to compare the differences between two independent samples when the sample distributions are not normally distributed and the sample sizes are small (n <30).

It is considered to be the nonparametric equivalent to the two-sample independent t-test." 
# https://www.statology.org/mann-whitney-u-test-r/ 

```{r}
HR <- read_excel("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Inputs/HomerangeOverview.xlsx")

# multiple by 100 to conver the area from km2 to hectares 
HR$hectares <- as.numeric(HR$HRarea*100)
temp <- HR %>% dplyr::select(BatID, HRarea, hectares)
temp

# check to see if this effects the mann whitney u test 

HR$Species <- as.factor(HR$Species)
HR$BatID <- as.factor(HR$BatID)
mbrahr <- HR %>% 
  filter(Species == "M.brandtii") %>% 
  droplevels() %>% 
  select(hectares) 
mbrahr <- mbrahr$hectares

mmyshr <- HR %>% 
  filter(Species == "M.mystacinus") %>% 
  droplevels() %>% 
  select(hectares) 
mmyshr <- mmyshr$hectares
# 

summary(mbrahr)
  # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  #  11.0   160.0   194.0   399.9   523.0  1274.0
sd(mbrahr)

summary(mmyshr)
   # Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   # 8.00   36.75   51.50   71.75  109.75  187.00
sd(mmyshr)

# 	Wilcoxon rank sum exact test
wilcox.test(mbrahr, mmyshr)
# still the same! 
# 
# data:  mbrahr and mmyshr
# W = 92, p-value = 0.00562
# alternative hypothesis: true location shift is not equal to 0
# There is strong evidence that the home range size between the two species is distinct from each other. 
```

## Now check the distance between available points for the bats locations 
```{r}
#Used and available positions - just for referencing. 
rsf.data <- read_csv("C:/Users/apmc/OneDrive - Norwegian University of Life Sciences/1. Nittedal 2018-2020/Nittedal_Main/Analyses/Outputs/HomeRangeExploration_2023-09-28/used_available_pts_combined95KE.csv", 
      col_types = cols(...1 = col_skip()))

rsfcoords <- coordinates(rsf.data) <- rsf.data[,c("Easting","Northing")]
#assigning coordinates 
library(sp)

proj4string(rsf.data)<-"+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
#Define the projection 
crs(rsf.data)

rsf.sf <- st_as_sf(rsf.data, crs = "+proj=utm + zone=32 + datum=WGS84 + units=m + no_defs")
# plot(rsf.sf )
# mapview(sMMYS) ## all good here

## Calculate nearest neighbors 
# Calculate the distance matrix
dist_matrix <- st_distance(rsf.sf)

# For each point, find the index of the nearest other point
# Exclude self (distance == 0)
nearest_index <- apply(dist_matrix, 1, function(x) {
  # Replace self-distance with Inf to exclude it from min
  x[x == 0] <- Inf
  which.min(x)
})

# Get the nearest distances
nearest_distance <- apply(dist_matrix, 1, function(x) {
  x[x != 0][which.min(x[x != 0])]
})

# Add results to your data
rsf.sf$nearest_id <- nearest_index
rsf.sf$nearest_dist <- nearest_distance

#print(points_sf)


## Now find the nearest neighbor, taking into consideration different individuals 
summary_stats <- rsf.sf %>% dplyr::filter(used == 0) %>% 
  group_by(BatID) %>%
  summarise(
    min_nn = min(nearest_dist, na.rm = TRUE),
    max_nn = max(nearest_dist, na.rm = TRUE),
    mean_nn = mean(nearest_dist, na.rm = TRUE)
  )

summary(summary_stats)
 #    BatID               min_nn            max_nn          mean_nn                geometry 
 # Length:21          Min.   :0.06774   Min.   : 14.76   Min.   : 3.881   MULTIPOINT   :21  
 # Class :character   1st Qu.:0.08153   1st Qu.: 24.48   1st Qu.: 7.086   epsg:NA      : 0  
 # Mode  :character   Median :0.24341   Median : 35.73   Median : 9.319   +proj=utm ...: 0  
 #                    Mean   :0.32303   Mean   : 59.73   Mean   :12.385                     
 #                    3rd Qu.:0.44954   3rd Qu.: 63.78   3rd Qu.:13.059                     
 #                    Max.   :1.05177   Max.   :254.08   Max.   :41.964 


```

